### 2023-02-21 figures/figures.Rmd
repo infra

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
- \usepackage{float}
- \usepackage{caption}
- \pagenumbering{gobble}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
knitr::opts_chunk$set(fig.pos = "H")
```

\pagebreak

```{r}
source("../scripts/99_utils.R")
```

```{r map}
print(2)
```

### 2023-02-21 scripts/99_utils.R

@@ -1,0 +1,5 @@
+# library(openair)
+# library(RFlux) # devtools::install_github("icos-etc/RFlux")
+# library(amerifluxr)
+
+print(1)

### 2023-02-21 figures/figures.Rmd
data loading

@@
 ```{r}
+if(dir.exists("../scripts")){
     source("../scripts/99_utils.R")
+}else{
+    source("scripts/99_utils.R")
+}
+
+dt <- read.csv(
+    "../../Data/Euroflux/BE-Lon/L2-L4_2004-2012/EFDC_L2_Flx_BELon_2008_v017_30m.txt",
+    header = TRUE,
+    colClasses = c("TIMESTAMP_START" = "character", "TIMESTAMP_END" = "character"), 
+    na.strings = "-9999") %>%
+    clean_names()
+
+head(dt)
 ```

### 2023-02-21 scripts/99_utils.R

@@ -1,5 +1,5 @@
+library(janitor)
 # library(openair)
 # library(RFlux) # devtools::install_github("icos-etc/RFlux")
 # library(amerifluxr)
 
-print(1)

### 2023-03-02 figures/figures.Rmd
clean up

--- 2023-02-21 figures/figures.Rmd
@@
 
+```{r, echo=FALSE}
+if(dir.exists("../scripts")){
+    shim <- "../"
+    source("../scripts/99_utils.R")
+}else{
+    shim <- ""
+    source("scripts/99_utils.R")
+}
+
+```
+
 ```{r}
-source("../scripts/99_utils.R")
+dt <- read.csv(
+    paste0(shim, "../../Data/Euroflux/BE-Lon/L2-L4_2004-2012/EFDC_L2_Flx_BELon_2008_v017_30m.txt"),
+    header = TRUE,
+    colClasses = c("TIMESTAMP_START" = "character", "TIMESTAMP_END" = "character"),
+    na.strings = "-9999") %>%
+    clean_names()
+
+head(dt)
 ```
@@
 ```{r map}
-print(2)
+
+coords_ire <- sf::st_as_sf(
+    data.frame(lat = 50.45055, lon = 4.5350415),
+    coords = c("lon", "lat"), crs = 4326)
+
+dt_sites <- read.csv(
+    system.file("extdata", "Site_metadata.csv", package = "FluxnetLSM")) %>%
+    clean_names() %>%
+    dplyr::mutate_all(na_if, "") %>%
+    st_as_sf(coords = c("site_longitude", "site_latitude"), crs = 4326)
+
+dt_dist <- unlist(st_distance(coords_ire, dt_sites))
+dt_sites_close <- dt_sites[dt_dist < units::as_units(100, "mi"),]
+
+rmarkdown::render_delayed({
+    mapview(dt_sites_close) + mapview(coords_ire, color = "red")
+    })
 ```

### 2023-03-02 scripts/99_utils.R

@@ -1,5 +1,8 @@
-library(janitor)
+suppressWarnings(suppressMessages(library(sf)))
+suppressWarnings(suppressMessages(library(dplyr)))
+suppressMessages(library(janitor))
+library(mapview)
+library(FluxnetLSM)
 # library(openair)
 # library(RFlux) # devtools::install_github("icos-etc/RFlux")
 # library(amerifluxr)
-

### 2023-03-06 figures/figures.Rmd
rm extra space
print distances

@@
 }
-
 ```
@@
 
-dt_dist <- unlist(st_distance(coords_ire, dt_sites))
-dt_sites_close <- dt_sites[dt_dist < units::as_units(100, "mi"),]
+dt_sites$dist <- unlist(list(unlist(st_distance(coords_ire, dt_sites)))) / 1609.34
+dt_sites_close <- dt_sites[dt_sites$dist < 100,]
+head(dplyr::arrange(dt_sites_close, dist)$dist)
 

### 2023-03-09 test.py
towards ffp with real data

@@ -1,0 +1,44 @@
+import os
+import janitor # clean_names
+import pandas as pd
+import seaborn as sns
+from pyffp import calc_footprint_FFP as myfootprint
+
+if os.path.exists("../scripts"):
+    shim = "../"
+else:
+    shim = ""
+
+# dt <- read.csv(
+#     paste0(shim, "../../Data/Euroflux/BE-Lon/L2-L4_2004-2012/EFDC_L2_Flx_BELon_2008_v017_30m.txt"),
+#     header = TRUE,
+#     colClasses = c("TIMESTAMP_START" = "character", "TIMESTAMP_END" = "character"),
+#     na.strings = "-9999") %>%
+#     clean_names()
+
+dt = pd.read_csv(
+    shim
+    + "../../Data/Euroflux/BE-Lon/L2-L4_2004-2012/EFDC_L2_Flx_BELon_2008_v017_30m.txt",
+    na_values=["-9999"],
+    dtype = {"TIMESTAMP_START":str, "TIMESTAMP_END":str}
+).clean_names()
+dt = dt[dt["ustar"] >= 0.1].reset_index(drop=True)
+
+# sns.histplot(data=dt, x="ustar")
+# plt.show()
+
+ustar = dt["ustar"].values[0]
+umean = dt["ws"].values[0]
+sigmav = dt["v_sigma"].values[0]
+wind_dir = dt["wd"].values[0]
+
+FFP = myfootprint.FFP(
+    zm=20.0,
+    umean=umean,
+    h=2000.0,
+    ol=-100.0,
+    sigmav=sigmav,
+    ustar=ustar,
+    wind_dir=wind_dir,
+    rs=[20.0, 40.0, 60.0, 80.0],
+)

### 2023-03-21 figures/figures.Rmd
look at more distant sites

@@
 dt_sites$dist <- unlist(list(unlist(st_distance(coords_ire, dt_sites)))) / 1609.34
-dt_sites_close <- dt_sites[dt_sites$dist < 100,]
-head(dplyr::arrange(dt_sites_close, dist)$dist)
+dt_sites_close <- dt_sites[dt_sites$dist < 100,] %>%
+    dplyr::arrange(dist)
+head(dt_sites_close$dist)
+st_coordinates(dt_sites_close[1,])
 

### 2023-03-21 test.py
geospatial viz with footprints

@@ -1,8 +1,17 @@
+# I think the climatology functions return an average over all the inputs not a time-series of footprints
+
 import os
 import janitor # clean_names
+import numpy as np
 import pandas as pd
+import geopandas as gpd
 import seaborn as sns
-from pyffp import calc_footprint_FFP as myfootprint
+import matplotlib.cm as cm
+import matplotlib.pyplot as plt
+from pyffp import calc_footprint_FFP_climatology as myfootprint
+
+from pyffp import utils as ffputils
+get_dd = ffputils.get_dd
 
 if os.path.exists("../scripts"):
     shim = "../"
@@ -23,22 +32,66 @@ dt = pd.read_csv(
     dtype = {"TIMESTAMP_START":str, "TIMESTAMP_END":str}
 ).clean_names()
 dt = dt[dt["ustar"] >= 0.1].reset_index(drop=True)
+dt["timestamp_start"] = pd.to_datetime(dt["timestamp_start"])
+dt["timestamp_end"] = pd.to_datetime(dt["timestamp_end"])
 
 # sns.histplot(data=dt, x="ustar")
 # plt.show()
 
-ustar = dt["ustar"].values[0]
-umean = dt["ws"].values[0]
-sigmav = dt["v_sigma"].values[0]
-wind_dir = dt["wd"].values[0]
+ustar = list(dt["ustar"].values[0:10])
+umean = list(dt["ws"].values[0:10])
+sigmav = list(dt["v_sigma"].values[0:10])
+wind_dir = list(dt["wd"].values[0:10])
 
-FFP = myfootprint.FFP(
-    zm=20.0,
-    umean=umean,
-    h=2000.0,
-    ol=-100.0,
+zm = [20.0 for i in range(10)]
+h = [2000.0 for i in range(10)]
+ol = [-100.0  for i in range(10)]
+
+FFP = myfootprint.FFP_climatology(
+    zm=zm,
+    z0=0.01,
+    # umean=umean,
+    h=h,
+    ol=ol,
     sigmav=sigmav,
     ustar=ustar,
-    wind_dir=wind_dir,
-    rs=[20.0, 40.0, 60.0, 80.0],
+    wind_dir=wind_dir
 )
+
+[k for k in FFP.keys()]
+len(FFP["rs"])
+len(FFP['fr'])
+len(FFP['xr'][0])
+len(FFP['yr'][0])
+
[…]

### 2023-03-21 figures/figures.Rmd
towards cartopy maps

--- 2023-03-06 figures/figures.Rmd
@@
 dt_sites$dist <- unlist(list(unlist(st_distance(coords_ire, dt_sites)))) / 1609.34
-dt_sites_close <- dt_sites[dt_sites$dist < 100,]
-head(dplyr::arrange(dt_sites_close, dist)$dist)
+dt_sites_close <- dt_sites[dt_sites$dist < 100,] %>%
+    dplyr::arrange(dist)
+head(dplyr::select(dt_sites_close, site_code, dist))
+
+st_coordinates(dt_sites_close[1,])
 

### 2023-03-21 figures/footprint.py

@@ -4,12 +4,14 @@ import os
 import janitor # clean_names
 import numpy as np
 import pandas as pd
-import geopandas as gpd
 import seaborn as sns
+import geopandas as gpd
+import cartopy.crs as ccrs
 import matplotlib.cm as cm
 import matplotlib.pyplot as plt
-from pyffp import calc_footprint_FFP_climatology as myfootprint
+import cartopy.feature as cfeature
 
+from pyffp import calc_footprint_FFP_climatology as myfootprint
 from pyffp import utils as ffputils
 get_dd = ffputils.get_dd
 
@@ -87,11 +89,17 @@ cc = [c]*len(levs)
 # generate GeoDataFrame of contours
 fig, ax = plt.subplots(figsize=(10, 8))
 cp = ax.contour(x_2d_dd, y_2d_dd, fs, levs, colors = cc, linewidths=0.5)
-gdf = ffputils.contour_to_gdf(cp, "test.gpkg")
-gdf["timestamp_start"] = list(dt["timestamp_start"].values[0:10])
+gdf = ffputils.contour_to_gdf(cp)
 # plt.show()
 plt.close()
 
-gpd.GeoDataFrame(geometry=gpd.points_from_xy(x=[origin_lon], y=[origin_lat])).to_file("BELon.gpkg", driver="GPKG")
-gdf.to_file("test.gpkg", driver="GPKG")
+be_lon = gpd.GeoDataFrame(geometry=gpd.points_from_xy(x=[origin_lon], y=[origin_lat]))
 
+fig = plt.figure(figsize=[6, 6])
+ax1 = fig.add_subplot(1, 1, 1, projection=ccrs.PlateCarree())
+ax1.add_feature(cfeature.LAND)
+gdf.plot(ax=ax1, legend=True, legend_kwds={"shrink": 0.6})
+be_lon.plot(ax=ax1, color="black", markersize=2)
+plt.tight_layout()
+# plt.show()
+plt.savefig("figures/footprint.pdf", bbox_inches="tight")

### 2023-03-21 scripts/kelsey.py
start kelsey reproduction

@@ -1,0 +1,31 @@
+import janitor # for .clean_names
+import pandas as pd
+import seaborn as sns
+
+dt = pd.read_csv(
+    "../../Data/Euroflux/BE-Lon/L2-L4_2004-2012/EFDC_L2_Flx_BELon_2008_v017_30m.txt"
+).clean_names()
+
+# --- define before, during, and after periods
+# 10 days before and after
+# Before: Aug 12, 13, 14, 15, 16, 17, 18, 19, 20, 21
+# During: Aug 22, 23, 24
+# After: Aug 25, 26, 27, 28, 29, 30, 31, 1,2,3
+dt["timestamp_start"] = pd.to_datetime(dt["timestamp_start"], format="%Y%m%d%H%M")
+dt["timestamp_end"] = pd.to_datetime(dt["timestamp_end"], format="%Y%m%d%H%M")
+dt["period"] = None
+dt.loc[
+    (dt["timestamp_end"] < pd.to_datetime("2008-08-22"))
+    & (dt["timestamp_start"] > pd.to_datetime("2008-08-11")),
+    "period",
+] = "before"
+dt.loc[
+    (dt["timestamp_end"] < pd.to_datetime("2008-08-25"))
+    & (dt["timestamp_start"] > pd.to_datetime("2008-08-21")),
+    "period",
+] = "during"
+dt.loc[dt["timestamp_start"] > pd.to_datetime("2008-08-24"),
+    "period",
+] = "after"
+
+dt[["timestamp_start", "timestamp_end", "period"]].tail(20)

### 2023-03-22 figures/figures.Rmd
label mapview pnts

@@
 ```{r map}
-
 coords_ire <- sf::st_as_sf(
@@
     coords = c("lon", "lat"), crs = 4326)
+coords_ire$site_code <- "Fleurus"
 
@@
 
-rmarkdown::render_delayed({
-    mapview(dt_sites_close) + mapview(coords_ire, color = "red")
-    })
+m1_data <- dplyr::bind_rows(dt_sites_close, coords_ire)
+m1 <- mapview(m1_data)
+addStaticLabels(m1, label = m1_data$site_code)
 ```

### 2023-03-22 scripts/99_utils.R

@@ -3,6 +3,7 @@ suppressWarnings(suppressMessages(library(dplyr)))
 suppressMessages(library(janitor))
 library(mapview)
 library(FluxnetLSM)
+library(leafem)
 # library(openair)
 # library(RFlux) # devtools::install_github("icos-etc/RFlux")
 # library(amerifluxr)

### 2024-04-15 manuscript/manuscript.tex
ms setup

i\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{listings}
\usepackage{lineno}
\usepackage{setspace}

\linenumbers

\title{Eddy covariance towers as sentinels of abnormal radioactive material releases}

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
\date{} 					% Or removing it

\author{
	\href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Jemma Stachelek},
	\href{https://orcid.org/0000}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}
	 Second Author},
	\href{https://orcid.org/0000}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}
	 Third Author} \\
	Earth and Environmental Sciences\\
	Los Alamos National Laboratory\\
	Los Alamos, NM, USA, 87544 \\
	\texttt{jsta@lanl.gov} \\
}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
\renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.000/XXXXX}}
\renewcommand{\undertitle}{This is a draft of a Manuscript in preparation for publication \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
\renewcommand{\shorttitle}{Eddy covariance towers and abnormal releases}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={Eddy covariance towers as sentinels of abnormal radioactive material releases},
pdfsubject={stat.AP},
pdfauthor={Jemma Stachelek}
}

\begin{document}
\maketitle
\doublespacing

\section*{Abstract}	
    Ensuring accurate detection and attribution of abnormal releases of radioactive material is critical for ensuring human health and safety. Most commonly, such detection is accomplished via active monitoring approaches involving the collection of physical samples. This is labor intensive and limits the temporal and spatial resolution of any detected events to a relatively coarse level. As an alternative, we developed a passive monitoring approach using eddy flux tower data records to identify signals from a known abnormal release and quantify the extent to which that signal also occurs at other times in the data record. We tested our approach relative to a known release in Belgium nominally occurring on August 23, 2008. With this approach, we identified several potential heretofore unidentified abnormal releases. We found that the occurrence of these releases was consistent with wind direction from the known release site to a given flux tower. We conclude that although eddy flux tower data records are a potentially rich target for these types of data mining efforts, geographic sparsity in tower locations as well as noisiness of the data streams may limit their usefulness to targeted constituents in relatively limited geographic regions.

\section{Introduction}
Detecting abnormal releases of radioactive material (hereafter simply “abnormal releases”) is important as a means of ensuring human health and industrial safety. The detection of such releases is typically accomplished using active approaches that require site-specific trace gas measurements \citep{carriganTraceGasEmissions1996} sometimes as a part of international monitoring networks (e.g., Sangiorgi et al. 2020). A potential shortcoming of these active monitoring approaches is that they require collection of physical samples (particulates, gasses, and other signal carriers) to retrospectively determine the nature of the release and the events that may have led up to it. Such active monitoring is thus labor intensive and limits the temporal and spatial resolution of any detected events to a relatively coarse level.

A potential alternative to active monitoring is passive monitoring (e.g., seismic monitoring for underground nuclear explosions, UNEs, Li et al. 2023). Although existing passive monitoring approaches are already critical components of abnormal release detection and tracking programs, diversification-of-approach may increase the programs' sensitivity and comprehensiveness. With a greater diversity of approaches, monitoring programs may be able to detect a wider array of constituents with differing characteristics. In the present study, we developed a novel passive monitoring approach to supplement existing programs that involves  the use of eddy flux tower data records to capture the signal from “known'' abnormal releases. We then use this signal to investigate the extent to which it also occurs at other times in the data record. We focused on eddy flux towers as they are a potential source of abnormal release “signatures,” are globally distributed, have been in near continuous operations for several decades, and can be relatively cheaply and easily deployed in locations of interest. Therefore, they constitute a rich target for data mining approaches aimed at identifying signatures resulting from direct interference of radioactive material with target measurements and/or the indirect signal of vegetation responses to the deposition of radioactive material.

Our use of flux tower records for signal detection stands in contrast to their typical use for direct carbon balance accounting. Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange (Baldocchi et al. 2018), our use treats each variable as a potential indirect signal of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc). In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation. We specifically focused on short-term vegetation stress responses (on the order of days to weeks) to potential deposition of radioactive material rather than chronic long-term signals (on the order of years to decades). First, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from a site in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008. The release originated from waste tanks into the atmosphere at the Institut des Radioelements (IRE) and lasted for several days totaling 50 GBq 131I. Grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg (Carle et al. 2010). We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers.

\section{Methods}

\subsection{Data Description}

We gathered measurements from eddy covariance platforms (i.e., “towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/). Only three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1). Apart from standard content formatting (e.g., timestamps, dropping empty columns, etc.), we processed each data file by 1) excluding nighttime measurements based on photosynthetic photon flux density (PPFD) or net solar radiation (netrad) and 2) defining an “event period” based on the publicly known date of the abnormal release. We focused on the following potential data streams in the eddy covariance tower records where the following were treated as dependent “response” variables: carbon dioxide, carbon dioxide flux (fc), latent heat (le), and sensible heat flux (h). Conversely, the following were treated as independent explanatory variables: wind speed, precipitation, atmospheric pressure, relative humidity, photosynthetic photon flux density, air temperature, and net radiation. We also used wind direction data from the eddy flux towers to determine when or if any released material might plausibly reach a particular location. Each of these variables were typically recorded by the flux tower instrumentation at a 30 minute interval barring any gaps due to missing data.

\subsection{Modeling approach}

Our analysis screened pairwise relationships between the aforementioned independent and dependent variables starting using a linear regression fit between each pair for the entire period of record. We excluded variable pairs from further analysis if the overall coefficient of determination (R2) was less than 0.1. We determined from visual inspection that many of the rejected variable pairs appeared to have a nonlinear relationship. We then fit an interaction model of the following form where Y represents a continuous response (e.g., CO2, latent heat, etc.), X is an independent variable (e.g., air temperature, humidity, etc.), and W is a categorical variable representing the “period” of observation (i.e., during, before, or after, the event):

% Y = 0+ 1X1 + B2W2+ 3X1W2

We fit this interaction model for a “window” of time encompassing the known release event and for every other possible window in the period-of-record. We treated the overall length of the window as an adjustable hyperparameter because we lacked information on exactly how long an abnormal material release might take to reach the footprint of the eddy covariance tower. We also lacked information on how long deposited material might take to induce a vegetation response. In the default base case, we set the window length (n\_days) as 7 days but we ran hyperparameter experiments to determine the effect of setting it to longer lengths of 10 and 14 days. We compared the “effect size” of the interaction term (B3) during the event-encompassing window against all other windows in a “rolling” analysis. Therefore, the overall analysis compares the interaction term of the window of the known event (August 23, 2008) against a window centered on every other observation. We “flagged” specific windows that had an effect size at least as large as that of the known event and with wind conditions at least as “towards” the flux tower as the event as “event detections”. We report the frequency of these event detections as “false positives”.

To accomplish this flagging of event detections (i.e., false positives), we defined several adjustable hyperparameters to deal with uncertainty surrounding atmospheric transport and the timing of any vegetation responses to material deposition. The first hyperparameter we defined, wind\_tolerance, reflects the range around the bearing from the facility to the tower location which counts as towards the tower. In the base case, we set wind\_tolerance to 10 degrees but we explored setting it to smaller values down to 5 degrees and larger values up to 20 degrees. The second hyperparameter we defined, event\_quantile\_effect, is an attempt to deal with the fact that we don't know the travel time of any materials to be deposited via an abnormal release or what if any delays exist between deposition and vegetation response. Rather than selecting the maximum value of B3 during a given window or simply the value at the exact time of the event, we set the value of event\_quantile\_effect to correspond with a specific quantile of all B3 values during the event window. In the base default case, this was set at 0.9 but we explored values as low as 0.5. All statistical analyses were carried out using the statsmodels Python package (Seabold and Perktold, 2009). Our processing scripts are openly available at [doi link], we refer readers to the original providers for data access.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{../figures/__map}
	\caption{Map of the three closest Fluxnet tower sites to the Fleurus release location. Note that BE-Lon and BE-Vie are of a similar easterly direction to the site whereas BE-Bra is of an entirely different northerly direction.}
	\label{fig:study_site}
\end{figure}

\section{Results}

\begin{figure}
	\centering
	\includegraphics[width=11cm]{../figures/__interaction_belon}
	\caption{Interaction plots at the BE-Lon site between air temperature (ta) and CO2 (co2) for the period of the known abnormal release (A) and an arbitrary non-release period (B). Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
	\label{fig:interaction}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=8cm]{../figures/__rolling_fleurus}
	\caption{Time series plots of rolling event detection analysis for the CO2 and air temperature (ta) variable pair. Panel labels indicate the name of the site, the variable pair (in parentheses), and the false positive rate (a). Event detection lines (orange), interaction effect size (blue), and event effect size (green) correspond to the left vertical axis. Fraction that the wind direction was towards the particular tower (black) corresponds to the right vertical axis.}
	\label{fig:rolling}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=12cm]{../figures/__hyperparameter_experiment}
	\caption{Sensitivity of false event detection to different hyperparameter values.}
	\label{fig:hyperparameter}
\end{figure}

\section{Discussion}

\section{Data Availability Statement}

All data and code associated with this manuscript will be made available by the time of publication (or earlier pending the conclusion of US Department of Energy approvals) with a Zenodo DOI at: \href{https://doi.org}{https://doi.org}

\section{Acknowledgements}

This are Acknowledgements

\bibliographystyle{jsta}
\bibliography{fluxnet}

\end{document}

### 2024-04-15 figures/maps.R
increase maps font size, coloring fix

@@ -54,7 +54,8 @@ gg_euro_overview <- ggmap(gg_map) +
   geom_text(
     data = m1_data, aes(x = X, y = Y, label = site_code),
     vjust = 0,
-    hjust = 0
+    hjust = 0,
+    size = 6
   )
 # gg_euro_overview
 
@@ -62,10 +63,10 @@ dt_sub <- dplyr::filter(m1_data, site_code == "BE-Lon")
 gg_euro_belon <- get_gg_sub(dt_sub)
 
 dt_sub <- dplyr::filter(m1_data, site_code == "BE-Vie")
-gg_euro_bevie <- get_gg_sub(dt_sub)
+gg_euro_bevie <- get_gg_sub(dt_sub, label_color = "white")
 
 dt_sub <- dplyr::filter(m1_data, site_code == "BE-Bra")
-gg_euro_bebra <- get_gg_sub(dt_sub)
+gg_euro_bebra <- get_gg_sub(dt_sub, label_color = "white")
 
 gg <- cowplot::plot_grid(gg_euro_overview, gg_euro_bebra, gg_euro_belon, gg_euro_bevie)
 ggsave("figures/__map.pdf", gg)

### 2024-04-15 scripts/99_utils.R

@@ -37,7 +37,7 @@ amf_clean <- function(fpath) {
   base1
 }
 
-get_gg_sub <- function(dt_sub, buffer_x = 0.02, buffer_y = 0.02) {
+get_gg_sub <- function(dt_sub, buffer_x = 0.02, buffer_y = 0.02, label_color = "black") {
   bbox_sub <- c(left   = dt_sub$X - buffer_x,
     bottom = dt_sub$Y - buffer_y,
     right  = dt_sub$X + buffer_x,
@@ -55,6 +55,8 @@ get_gg_sub <- function(dt_sub, buffer_x = 0.02, buffer_y = 0.02) {
       alpha = .8) +
     geom_text(aes(x = X, y = Y, label = site_code), data = dt_sub,
       vjust = 0,
-      hjust = 0)
+      hjust = 0,
+      size = 6,
+      color = label_color)
   res
 }

### 2024-04-15 figures/interaction.py
add units to y-axis co2

@@ -10,6 +10,8 @@ from matplotlib.lines import Line2D
 sys.path.append(".")
 from src import rolling
 
+plt.rcParams["text.usetex"] = True
+
 
 def hue_regplot(data, x, y, hue, palette=None, **kwargs):
     """
@@ -91,6 +93,7 @@ legend_elements = [
     Line2D([0], [0], color=colors[2], lw=4, label="After"),
 ]
 ax1.legend(handles=legend_elements, loc="upper right")
+ax1.set_ylabel("$CO_2$ (ppm)")
 ax1.set_title("A.", pad=0, x=0.07, y=0.9)
 #
 gg2 = sns.histplot(
@@ -110,6 +113,7 @@ ts_start = timestamps[
 ]
 dt_event_weak = rolling.define_period(dt_select, n_days=n_days, date_event=ts_start)
 gg3 = hue_regplot(data=dt_event_weak, x="ta", y="co2", hue="period", ax=ax3, ci=None)
+ax1.set_ylabel("$CO_2$ (ppm)")
 ax3.set_title("B.", pad=0, x=0.07, y=0.9)
 #
 # stats.percentileofscore(pfdist["fdist"], fevent["fevent"], nan_policy="omit") / 100

### 2024-04-15 figures/interaction.py
limit comparison panel to same month as real event

@@ -107,13 +107,20 @@ gg2.axvline(fevent["fevent"].values[0], color="orange")
 ax2.set_xlabel("Effect size")
 ax2.set_title("C.", pad=0, x=0.07, y=0.9)
 #
-timestamps = [x for x in pfdist["timestamp"]]
-ts_start = timestamps[
-    int(np.where([x == np.nanmin(pfdist["fdist"]) for x in pfdist["fdist"]])[0][0])
-]
+
+# limit ts_start to the same month as date_event
+pfdist["month"] = [x.month for x in pd.to_datetime(pfdist["timestamp"])]
+# avoid 2009 because of missing data
+pfdist["year"] = [x.year for x in pd.to_datetime(pfdist["timestamp"])]
+pfdist_temp = pfdist[pfdist["year"] != 2009]
+pfdist_temp = pfdist_temp[pfdist_temp["year"] != 2008]
+pfdist_temp = pfdist_temp.sort_values("fdist", ascending=True)
+pfdist_temp = pfdist_temp[pfdist_temp["month"] == pd.to_datetime(date_event).month]
+ts_start = pfdist_temp["timestamp"].iloc[[0]].values[0]
+
 dt_event_weak = rolling.define_period(dt_select, n_days=n_days, date_event=ts_start)
 gg3 = hue_regplot(data=dt_event_weak, x="ta", y="co2", hue="period", ax=ax3, ci=None)
-ax1.set_ylabel("$CO_2$ (ppm)")
+ax3.set_ylabel("$CO_2$ (ppm)")
 ax3.set_title("B.", pad=0, x=0.07, y=0.9)
 #
 # stats.percentileofscore(pfdist["fdist"], fevent["fevent"], nan_policy="omit") / 100

### 2024-05-24 figures/figures.Rmd
linting

@@
 ```{r, echo=FALSE}
-if(dir.exists("../scripts")){
+if (dir.exists("../scripts")) {
   shim <- "../"
   source("../scripts/99_utils.R")
-}else{
+} else {
   shim <- ""
@@
 dt_sites$dist <- unlist(list(unlist(st_distance(coords_ire, dt_sites)))) / 1609.34
-dt_sites_close <- dt_sites[dt_sites$dist < 100,] %>%
+dt_sites_close <- dt_sites[dt_sites$dist < 100, ] %>%
   dplyr::arrange(dist)
@@
 
-st_coordinates(dt_sites_close[1,])
+st_coordinates(dt_sites_close[1, ])
 

### 2024-05-24 figures/maps.R
print site distance

@@ -17,7 +17,8 @@ dt_sites <- read.csv(
   # dplyr::mutate_all(na_if, "") %>%
   st_as_sf(coords = c("site_longitude", "site_latitude"), crs = 4326)
 
-dt_sites$dist <- unlist(list(unlist(st_distance(coords_ire, dt_sites)))) / 1609.34
+dt_sites$dist <- unlist(list(unlist(
+  st_distance(coords_ire, dt_sites)))) / 1609.34 # m to miles
 dt_sites_close <- dt_sites[dt_sites$dist < 100, ] %>%
   dplyr::arrange(dist)
 sf::st_write(dt_sites_close, "dt_sites_close.gpkg", append = FALSE)
@@ -27,11 +28,14 @@ sf::st_write(dt_sites_close, "dt_sites_close.gpkg", append = FALSE)
 
 m1_data <- dplyr::bind_rows(dt_sites_close, coords_ire)
 m1_data <- cbind(m1_data, st_coordinates(m1_data))
-m1_data <- sf::st_drop_geometry(dplyr::select(m1_data, X, Y, site_code))
 m1_data <- dplyr::filter(
   m1_data,
   site_code %in% c("BE-Lon", "BE-Vie", "BE-Bra", "IRE")
 )
+m1_data <- dplyr::mutate(m1_data, dist = dist * 1.60934)  # mi to km
+m1_data <- sf::st_drop_geometry(
+  dplyr::select(m1_data, X, Y, site_code, dist))
+
 
 print(m1_data)
 buffer_x <- 0.5

### 2024-08-12 scripts/00_get_ameriflux.R
add additional fukushima site

@@ -50,18 +50,26 @@ site_filtered <- site %>%
   dplyr::filter(data_start < 2002)
 
 head(site_filtered)
-res <- lapply(site_filtered$site_id, check_site)
+if (!file.exists("../../Data/Ameriflux/US-Wrc.csv")) {
+  res <- lapply(site_filtered$site_id, check_site)
 
-covers_fukushima <- as.logical(unlist(lapply(res, function(x) x[1])))
-res_out <- site_filtered[covers_fukushima, ]
-res_out$path <- as.character(unlist(lapply(res, function(x) x[2])))[covers_fukushima]
+  covers_fukushima <- as.logical(unlist(lapply(res, function(x) x[1])))
+  res_out <- site_filtered[covers_fukushima, ]
+  res_out$path <- as.character(unlist(lapply(res, function(x) x[2])))[covers_fukushima]
 
-write.csv(res_out, "data/ameriflux_pnw.csv", row.names = FALSE)
+  write.csv(res_out, "data/ameriflux_pnw.csv", row.names = FALSE)
+}
+
+# --- Interior US
+if (!file.exists("../../Data/Ameriflux/US-GLE.csv")) {
+  dt <- get_sites(41.298056, -106.1375, "test")
+  res <- check_site("US-GLE")
+}
 
 # --- Australia
 # dt <- get_sites(-33.867778, 151.21, "test")
 
 # --- Brazil
-test <- site[substring(site$SITE_ID, 0, 1) == "B", ]
-test <- test[test$DATA_POLICY == "CCBY4.0", ]
+# test <- site[substring(site$SITE_ID, 0, 1) == "B", ]
+# test <- test[test$DATA_POLICY == "CCBY4.0", ]
 # Brazil data doesn't span the required period :(

### 2024-08-12 scripts/01_fit_rolling.py

@@ -275,8 +275,8 @@ def fit_rolling(
         g2 = sns.lineplot(
             data=g_data, x="timestamp", y="wind_fraction", ax=ax2, color="black"
         )
-        g2.set_ylim(0, 0.5)
-        g2.set_yticks([0, 0.2, 0.4])
+        g2.set_ylim(0, 1)
+        g2.set_yticks([0, 0.2, 0.4, 0.6, 0.8])
 
         ax2.set_xlim([datetime(panel_start_year, 1, 1), datetime(panel_end_year, 2, 1)])
         # g_data.iloc[int(event_index)]["p"]
@@ -307,7 +307,7 @@ def fit_rolling(
             # + "="
             # + str(false_positive_rate),
             y=0.87,
-            x=0.175,
+            x=0.18,
         )
         ax1.set_xlabel("")
         ax2.set_xlabel("")

### 2024-08-12 scripts/01_fit_rolling.py
better grid formatting

@@ -247,7 +247,7 @@ def fit_rolling(
 
         # g_data["timestamp"] = g_data["timestamp"].replace({np.nan: None})
 
-        _, axs = plt.subplots(figsize=(9, 6), nrows=2)
+        _, axs = plt.subplots(figsize=(4, 6), nrows=2)
         ax1 = axs[0]
         ax2 = axs[1]
 
@@ -297,7 +297,9 @@ def fit_rolling(
             ax2.yaxis.label.set_color("white")
             plt.yticks(color="white")
         ax2.set_ylabel("fraction wind towards")
-        ax2.xaxis.set_major_formatter(mdates.DateFormatter("%Y"))
+        ax2.xaxis.set_major_locator(mdates.YearLocator())
+        # ax2.tick_params(axis="x", labelrotation=90)
+        ax2.xaxis.set_major_formatter(mdates.DateFormatter("%y"))
         plt.suptitle(
             site,
             # + "("
@@ -307,7 +309,7 @@ def fit_rolling(
             # + "="
             # + str(false_positive_rate),
             y=0.87,
-            x=0.18,
+            x=0.25,
         )
         ax1.set_xlabel("")
         ax2.set_xlabel("")

### 2024-08-13 scripts/01_fit_rolling.py
improved panelization

@@ -57,6 +57,7 @@ from src import rolling
 @click.option("--panel_ylim", type=int, default=360)
 @click.option("--panel_start_year", type=int, default=2004)
 @click.option("--panel_end_year", type=int, default=2013)
+@click.option("--wind_ylim", type=float, default=1.0)
 def fit_rolling(
     site,
     date_event,
@@ -80,6 +81,7 @@ def fit_rolling(
     panel_ylim=360,
     panel_start_year=2004,
     panel_end_year=2013,
+    wind_ylim=1,
 ):
     # --- setup
     dep_cols = ["co2", "fc", "le", "h", "co"]
@@ -275,8 +277,12 @@ def fit_rolling(
         g2 = sns.lineplot(
             data=g_data, x="timestamp", y="wind_fraction", ax=ax2, color="black"
         )
-        g2.set_ylim(0, 1)
-        g2.set_yticks([0, 0.2, 0.4, 0.6, 0.8])
+        g2.set_ylim(0, wind_ylim)
+        y_tick_range = [0, 0.2, 0.4, 0.6, 0.8]
+        y_tick_range = list(
+            itertools.compress(y_tick_range, [x <= wind_ylim for x in y_tick_range])
+        )
+        g2.set_yticks(y_tick_range)
 
         ax2.set_xlim([datetime(panel_start_year, 1, 1), datetime(panel_end_year, 2, 1)])
         # g_data.iloc[int(event_index)]["p"]
@@ -317,7 +323,7 @@ def fit_rolling(
         print(path_fig)
         plt.subplots_adjust(hspace=0.1)
         # plt.subplots_adjust(bottom=0, right=0.02, left=0, top=0.02)
-        plt.savefig(path_fig)
+        plt.savefig(path_fig, bbox_inches="tight")
 
     # --- save logging info
     log_info = pd.DataFrame(

### 2024-08-13 scripts/01_fit_rolling.py
no data masking

@@ -58,6 +58,8 @@ from src import rolling
 @click.option("--panel_start_year", type=int, default=2004)
 @click.option("--panel_end_year", type=int, default=2013)
 @click.option("--wind_ylim", type=float, default=1.0)
+@click.option("--mask_start", type=str, default=None)
+@click.option("--mask_end", type=str, default=None)
 def fit_rolling(
     site,
     date_event,
@@ -82,6 +84,8 @@ def fit_rolling(
     panel_start_year=2004,
     panel_end_year=2013,
     wind_ylim=1,
+    mask_start="2009-01-01",
+    mask_end="2010-01-01",
 ):
     # --- setup
     dep_cols = ["co2", "fc", "le", "h", "co"]
@@ -203,6 +207,28 @@ def fit_rolling(
         }
     )
     g_data["timestamp"] = pd.to_datetime(g_data["timestamp"])
+    # fill-in missing timestamps
+    nonnat_ts = g_data["timestamp"][~np.isnat(g_data["timestamp"].values)]
+    full_ts = pd.to_datetime(
+        np.arange(
+            min(nonnat_ts),
+            max(nonnat_ts),
+            np.timedelta64(30, "m"),
+            dtype="datetime64[m]",
+        )
+    )
+    missing_ts = list(set(full_ts) - set(g_data["timestamp"]))
+    g_data_missing_ts = pd.DataFrame(
+        {
+            "timestamp": missing_ts,
+            "index": pd.NA,
+            "p": pd.NA,
+            "wind_fraction": pd.NA,
+            "F": pd.NA,
+        }
+    )
+    g_data = pd.concat([g_data, g_data_missing_ts], ignore_index=True)
+
     if event_wind is None:
         event_wind = np.quantile(
             [
@@ -286,6 +312,40 @@ def fit_rolling(
 
         ax2.set_xlim([datetime(panel_start_year, 1, 1), datetime(panel_end_year, 2, 1)])
         # g_data.iloc[int(event_index)]["p"]
+
+        # mask out missing data with white bars
+        if mask_start is not None and mask_end is not None:
+            # mask_start = pd.to_datetime("2009-02-01")
+            # mask_end = pd.to_datetime("2009-12-01")
+            missing_mask = [
+                x and y
+                for x, y in zip(
+                    g_data["timestamp"] > mask_start, g_data["timestamp"] < mask_end
+                )
+            ]
+            # missing_mask = [
+            #     x and y and w and p
[…]

### 2024-08-14 figures/maps.R
toward fukushima map panels

@@ -52,7 +52,7 @@ gg_euro_bebra <- get_gg_sub(dt_sub, label_color = "white")
 gg <- cowplot::plot_grid(gg_overview, gg_euro_bebra, gg_euro_belon, gg_euro_bevie)
 ggsave("figures/__map_fleurus.pdf", gg)
 
-# --- fukushima
+# --- jp-fhk
 sites <- get_sites(37.423056, 141.033056, "Fukushima", 190)
 overview <- generate_overview(sites, c("JP-FHK", "Fukushima"),
   nudge_x = 0.6, label_color = "red", buffer_x = 0.7)
@@ -63,9 +63,9 @@ dt_sub <- dplyr::filter(m1_data, site_code == "JP-FHK")
 gg_jp_fhk <- get_gg_sub(dt_sub, label_color = "white")
 
 gg <- cowplot::plot_grid(gg_overview, gg_jp_fhk)
-ggsave("figures/__map_fukushima.pdf", gg)
+ggsave("figures/__map_fhk.pdf", gg)
 
-# ---
+# --- us-wrc
 sites <- get_sites(45.8205, -121.9519, "US-Wrc", 190)
 overview <- generate_overview(sites, c("US-Wrc"), buffer_x = 3, buffer_y = 2.5)
 gg_overview <- overview$gg_overview
@@ -76,3 +76,46 @@ gg_us_wrc <- get_gg_sub(dt_sub, label_color = "white")
 
 gg <- cowplot::plot_grid(gg_overview, gg_us_wrc)
 ggsave("figures/__map_wrc.pdf", gg)
+
+# -- us-gle
+sites <- get_sites(41.3665, -106.2399, "US-Gle", 190)
+overview <- generate_overview(sites, c("US-Gle"), buffer_x = 3, buffer_y = 2.5)
+gg_overview <- overview$gg_overview
+m1_data <- overview$m1_data
+
+dt_sub <- dplyr::filter(m1_data, site_code == "US-Gle" & is.na(dist))
+gg_us_gle <- get_gg_sub(dt_sub, label_color = "white")
+
+gg <- cowplot::plot_grid(gg_overview, gg_us_gle)
+ggsave("figures/__map_gle.pdf", gg)
+
+# -- oz-mul
+sites <- get_sites(-22.2828, 133.2493, "OZ-Mul", 190)
+overview <- generate_overview(sites, c("OZ-Mul"), buffer_x = 3, buffer_y = 2.5)
[…]

### 2024-08-15 figures/maps.R
assembled map panels

@@ -102,6 +102,7 @@ gg <- cowplot::plot_grid(gg_overview, gg_oz_mul)
 ggsave("figures/__map_mul.pdf", gg)
 
 # --- fukushima
+# https://r-spatial.org/r/2018/10/25/ggplot2-sf.html
 crsrobin <- "+proj=robin +lon_0=-180 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
 
 dt_fuku_df <- data.frame(
@@ -111,11 +112,15 @@ dt_fuku_df <- data.frame(
 dt_fuku <- st_transform(st_as_sf(dt_fuku_df, coords = c("lon", "lat"), crs = 4326), crs = crsrobin)
 dt_fuku_df <- cbind(dt_fuku_df, st_coordinates(dt_fuku))
 
+library(rnaturalearth)
 world <- ne_countries(scale = "medium", returnclass = "sf")
-gg <- ggplot(data = st_transform(st_break_antimeridian(world, lon_0 = 180), crs = crsrobin)) +
+gg_overview <- ggplot(data = st_transform(st_break_antimeridian(world, lon_0 = 180), crs = crsrobin)) +
   geom_sf() +
   geom_sf(data = dt_fuku) +
-  geom_text(data = dt_fuku_df, aes(x = X, y = Y, label = site_code),
+  geom_text(data = dt_fuku_df, aes(x = X, y = Y * 1.06, label = site_code),
     color = "darkblue", fontface = "bold", check_overlap = FALSE) +
-  theme_bw()
-ggsave("test.pdf", gg)
+  coord_sf(xlim = c(-4320952 * 1.3, 6376465  * 1.2), ylim = c(-2383193 * 1.3, 4891261 * 1.3), expand = FALSE) +
+  theme_bw() +
+  theme(axis.title = element_blank())
+gg <- cowplot::plot_grid(gg_overview, gg_us_wrc, gg_oz_mul, gg_us_gle)
+ggsave("figures/__map_fukushima.pdf", gg)

### 2024-08-15 manuscript/manuscript.tex
stash conversion progress

@@
-i\documentclass{article}
+\documentclass{article}
 
@@
 \author{
-	\href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Jemma Stachelek},
-	\href{https://orcid.org/0000}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}
-	 Second Author},
-	\href{https://orcid.org/0000}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}
-	 Third Author} \\
-	Earth and Environmental Sciences\\
+	\href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Jemma Stachelek$^1$} \\
+	$^1$ Earth and Environmental Sciences\\
 	Los Alamos National Laboratory\\
@@
 	\texttt{jsta@lanl.gov} \\
+	\And
+	Vachel A. Kraklow$^2$ \\
+	$^2$ Earth and Environmental Sciences\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\
+	\And
+	E. Christi Thompson$^3$ \\
+	$^3$ Earth and Environmental Sciences\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\
+	\And
+	\textbf{L. Turin Dickman}$^2$ \\
+	$^2$ Earth and Environmental Sciences\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\
+	\And
+	\textbf{Emily Casleton}$^3$ \\
+	$^3$ Earth and Environmental Sciences\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\
+	\And
+	\textbf{Sanna Sevanto}$^2$ \\
+	$^2$ Earth and Environmental Sciences\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\
+	\And
+	\textbf{Ann Junghans}$^4$ \\
+	$^4$ Earth and Environmental Sciences\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544
 }
@@
 \renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.000/XXXXX}}
-\renewcommand{\undertitle}{This is a draft of a Manuscript in preparation for publication \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
-\renewcommand{\shorttitle}{Eddy covariance towers and abnormal releases}
+\renewcommand{\undertitle}{This is a draft of a Manuscript in preparation for publication in Nature: Communications \\ \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
+\renewcommand{\shorttitle}{Eddy covariance towers as sentinels}
 
@@
 
+\vspace{-2em}
+
 \section*{Abstract}	
-    Ensuring accurate detection and attribution of abnormal releases of radioactive material is critical for ensuring human health and safety.
+Ensuring accurate detection and attribution of abnormal releases of radioactive material is critical for protecting human health and safety.
 Most commonly, such detection is accomplished via active monitoring approaches involving the collection of physical samples.
 This is labor intensive and limits the temporal and spatial resolution of any detected events to a relatively coarse level.
-As an alternative, we developed a passive monitoring approach using eddy flux tower data records to identify signals from a known abnormal release and quantify the extent to which that signal also occurs at other times in the data record.
-We tested our approach relative to a known release in Belgium nominally occurring on August 23, 2008.
-With this approach, we identified several potential heretofore unidentified abnormal releases.
-We found that the occurrence of these releases was consistent with wind direction from the known release site to a given flux tower.
-We conclude that although eddy flux tower data records are a potentially rich target for these types of data mining efforts, geographic sparsity in tower locations as well as noisiness of the data streams may limit their usefulness to targeted constituents in relatively limited geographic regions.
+As an alternative first step towards passive monitoring, we developed an approach using eddy flux tower data records to identify signals from a known abnormal release and quantify the extent to which that signal also occurs at other times in the data record.
+Through two case-studies, one of which targeted the Fukushima nuclear disaster and the other targeting a radioisotope production facility, we tested our approach and identified several potential heretofore unidentified abnormal releases that were consistent with atmospheric circulation patterns and/or wind direction from known release sites.
+Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
 
@@
 Detecting abnormal releases of radioactive material (hereafter simply “abnormal releases”) is important as a means of ensuring human health and industrial safety.
-The detection of such releases is typically accomplished using active approaches that require site-specific trace gas measurements \citep{carriganTraceGasEmissions1996} sometimes as a part of international monitoring networks (e.g., Sangiorgi et al.
+The detection of such releases is typically accomplished using active approaches that require site-specific sampling or trace gas measurements (Loyalka 1983) sometimes as a part of international monitoring networks (e.g., Sangiorgi et al.
 2020).
 A potential shortcoming of these active monitoring approaches is that they require collection of physical samples (particulates, gasses, and other signal carriers) to retrospectively determine the nature of the release and the events that may have led up to it.
-Such active monitoring is thus labor intensive and limits the temporal and spatial resolution of any detected events to a relatively coarse level.
+Active monitoring with such means is thus labor intensive and limits the temporal and spatial resolution of any detected events to a relatively coarse level.
 
-A potential alternative to active monitoring is passive monitoring (e.g., seismic monitoring for underground nuclear explosions, UNEs, Li et al.
-2023).
-Although existing passive monitoring approaches are already critical components of abnormal release detection and tracking programs, diversification-of-approach may increase the programs' sensitivity and comprehensiveness.
-With a greater diversity of approaches, monitoring programs may be able to detect a wider array of constituents with differing characteristics.
-In the present study, we developed a novel passive monitoring approach to supplement existing programs that involves  the use of eddy flux tower data records to capture the signal from “known'' abnormal releases.
-We then use this signal to investigate the extent to which it also occurs at other times in the data record.
-We focused on eddy flux towers as they are a potential source of abnormal release “signatures,” are globally distributed, have been in near continuous operations for several decades, and can be relatively cheaply and easily deployed in locations of interest.
+A potential alternative to active monitoring is passive monitoring (e.g., the German Integrated Measuring and Information system [IMIS] and the International Monitoring System [IMS], Bieringer and Schlosser, 2004).
+Although existing wide area monitoring systems are already critical components of abnormal release detection and tracking programs (Medici et al.
+2011), diversification-of-approach may increase the programs’ sensitivity and comprehensiveness.
+With a greater diversity of methods, monitoring programs may be able to detect a wider array of constituents with differing characteristics.
+In the present study, we developed a novel passive monitoring approach to supplement existing programs that involves the use of eddy covariance data to capture a signal from known abnormal releases.
+This signal is then used to investigate whether additional potential abnormal releases exist at other times in the data record.
+We focused on eddy flux towers, as they are a potential source of abnormal release signatures, are globally distributed, have been in near continuous operation for several decades, and can be relatively cheaply and easily deployed in locations of interest.
 Therefore, they constitute a rich target for data mining approaches aimed at identifying signatures resulting from direct interference of radioactive material with target measurements and/or the indirect signal of vegetation responses to the deposition of radioactive material.
 
-Our use of flux tower records for signal detection stands in contrast to their typical use for direct carbon balance accounting.
+Our utilization of flux tower records for event discrimination stands in contrast to the typical use of data from the towers for direct carbon balance accounting.
 Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange (Baldocchi et al.
-2018), our use treats each variable as a potential indirect signal of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc).
-In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation.
-We specifically focused on short-term vegetation stress responses (on the order of days to weeks) to potential deposition of radioactive material rather than chronic long-term signals (on the order of years to decades).
-First, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from a site in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008.
-The release originated from waste tanks into the atmosphere at the Institut des Radioelements (IRE) and lasted for several days totaling 50 GBq 131I.
-Grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg (Carle et al.
+2018), our use treats each variable as potentially including an indirect signal (i.e.
+an anomaly from normal vegetation behavior) of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc).
+In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation in a known way (Nobel 2020).
+
+We specifically focused on short-term responses of vegetation (on the order of days to weeks) to potential deposition of radioactive material rather than chronic long-term signals (on the order of years to decades).
+In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008.
+The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq 131I (Carle et al.
 2010).
+For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg.
+In the second case study, we further explored potential signals influx tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days eventually dispersing throughout the entire northern hemisphere (Meszaros et al.
+2016).
 We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers.
+While little is known about the direct or indirect effects of radioactive material on vegetation, we expect that exposure may cause declines in photosynthetic CO2 uptake.
+This may be expressed as a change in the slope of the CO2 relationship with other environmental variables (e.g.
+air temperature, humidity, precipitation, etc.) before, during, and after exposure.
 
@@
 
-We gathered measurements from eddy covariance platforms (i.e., “towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/).
+For the IRE case study, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/).
 Only three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1).
-Apart from standard content formatting (e.g., timestamps, dropping empty columns, etc.), we processed each data file by 1) excluding nighttime measurements based on photosynthetic photon flux density (PPFD) or net solar radiation (netrad) and 2) defining an “event period” based on the publicly known date of the abnormal release.
-We focused on the following potential data streams in the eddy covariance tower records where the following were treated as dependent “response” variables: carbon dioxide, carbon dioxide flux (fc), latent heat (le), and sensible heat flux (h).
+Each of the three flux tower sites are located within a crop ecosystem type and had almost 10 years of data available from 2004 through 2013 recorded at 30 min intervals.
+While BE-Lon had a near-continuous data record during this period, BE-Bra had a long period of missing data in 2005, and both BE-Bra and BE-Vie had a long period of missing data in 2009.
+The distance between IRE and the flux tower sites are approximately 19 km, 95 km, and 105 km for the BE-Lon, BE-Bra, and BE-Vie respectively.
+
+For the Fukushima case study, we gathered measurements from flux towers located in the Northern Hemisphere (Western United States), which are part of the Ameriflux network (https://ameriflux.lbl.gov/).
+For contrast to the Northern Hemisphere sites where we expect a higher likelihood of impact, we also gathered measurements from flux towers located in the Southern Hemisphere, which are part of the OzFlux network (Cleverly 2011).
+We selected sites with sufficient long-term data to cover the 2011 date of the disaster and excluded boreal sites with extreme snow cover related seasonality.
+We ultimately selected three sites (US-Wrc, US-GLE, and OZ-Mul, Figure 1B).
+
+\subsection{Modeling Approach}
+We processed each data file by 1) excluding nighttime measurements based on photosynthetic photon flux density (PPFD > 100 µmol m-2 s-1) or net solar radiation (netrad > 0 W/m2) and 2) defining an “event period” following the publicly released date of the abnormal release.
+We focused on the following variables as potentially containing a signature of the release : carbon dioxide, carbon dioxide flux (fc), latent heat (le), and sensible heat flux (h).
 Conversely, the following were treated as independent explanatory variables: wind speed, precipitation, atmospheric pressure, relative humidity, photosynthetic photon flux density, air temperature, and net radiation.
-We also used wind direction data from the eddy flux towers to determine when or if any released material might plausibly reach a particular location.
-Each of these variables were typically recorded by the flux tower instrumentation at a 30 minute interval barring any gaps due to missing data.
+Additionally, we used wind direction data from the eddy flux towers to determine when or if any released material might plausibly reach a particular location.
+Each of these variables were recorded by the flux tower instrumentation at a 30 minute (min) interval barring any gaps due to missing data.
 
-\subsection{Modeling approach}
+To begin our analysis, we screened all pairwise, linear relationships between the aforementioned independent and dependent variables using a linear regression fit between each pair for the entire period of record (n=51).
+Variable pairs were excluded from further analysis if the overall coefficient of determination ($R^2$) was less than 0.1.
+This low threshold was chosen to maximize the number of pairs that could potentially be investigated more deeply.
+From visual inspection, we determined that many of the rejected variable pairs appeared to have a nonlinear relationship.
+As our aim was to develop a methodology that could be applied broadly to other sites, we did not apply transformation to linearize these relationships without a physical justification for the transformation.
+Evaluating an exhaustive list of data transformations was beyond the scope of the present effort.
+After identification of candidate variable pairs, we fit an interaction model of the following form where Y represents a continuous response (e.g., $CO_2$, latent heat, etc.), X1 is an independent variable (e.g., air temperature, humidity, etc.), W2 is a categorical variable representing the “period” of observation (i.e., during, before, or after, the event), $\beta$0 is the intercept, $\beta$1,2 are slope parameters for X1 and W2 respectively, and $\beta$3 is the slope parameter for the interaction between X1 and W2:
+Y = 0+ 1X1 + B2W2+ 3X1W2
 
-Our analysis screened pairwise relationships between the aforementioned independent and dependent variables starting using a linear regression fit between each pair for the entire period of record.
-We excluded variable pairs from further analysis if the overall coefficient of determination (R2) was less than 0.1.
-We determined from visual inspection that many of the rejected variable pairs appeared to have a nonlinear relationship.
-We then fit an interaction model of the following form where Y represents a continuous response (e.g., CO2, latent heat, etc.), X is an independent variable (e.g., air temperature, humidity, etc.), and W is a categorical variable representing the “period” of observation (i.e., during, before, or after, the event):
+This interaction model is fit for a window of time encompassing the known release event and for every other possible window in the period-of-record.
+The overall length of the window was treated as an adjustable hyperparameter since the length of time for an abnormal material release to reach the footprint of the flux tower may be event-dependent and/or unknown.
+In addition, the length of time for deposited material to induce a possible vegetation response (or at least a flux tower anomaly) is an unknown quantity.
+Thus, in the default base case, we set the window length (n\_days) as 7 days, but we ran hyperparameter experiments to determine the result of setting it to longer periods of 10 and 14 days.
+The window length affects the duration of time designated as before and after the event.
+The period of time designated as during the event was set as 2 days in all cases.
+We compared the effect size of the interaction term ($\beta$3) during the event-encompassing window against all other windows in a “rolling” analysis.
+Therefore, the overall analysis compares the interaction term of the window of the known event (which occurred on August 23, 2008) against a window centered on every other observation.
+Specific windows that had an effect size at least as large as that of the known event and with wind conditions defined as “towards” the flux tower were flagged as “event detections”.
+The frequency of these event detections are reported as “event detections”.
 
-% Y = 0+ 1X1 + B2W2+ 3X1W2
-
-We fit this interaction model for a “window” of time encompassing the known release event and for every other possible window in the period-of-record.
-We treated the overall length of the window as an adjustable hyperparameter because we lacked information on exactly how long an abnormal material release might take to reach the footprint of the eddy covariance tower.
-We also lacked information on how long deposited material might take to induce a vegetation response.
-In the default base case, we set the window length (n\_days) as 7 days but we ran hyperparameter experiments to determine the effect of setting it to longer lengths of 10 and 14 days.
-We compared the “effect size” of the interaction term (B3) during the event-encompassing window against all other windows in a “rolling” analysis.
-Therefore, the overall analysis compares the interaction term of the window of the known event (August 23, 2008) against a window centered on every other observation.
-We “flagged” specific windows that had an effect size at least as large as that of the known event and with wind conditions at least as “towards” the flux tower as the event as “event detections”.
-We report the frequency of these event detections as “false positives”.
-
-To accomplish this flagging of event detections (i.e., false positives), we defined several adjustable hyperparameters to deal with uncertainty surrounding atmospheric transport and the timing of any vegetation responses to material deposition.
+In order to flag event detections, we defined several adjustable hyperparameters to deal with uncertainty surrounding atmospheric transport and the timing of any possible responses to material deposition.
 The first hyperparameter we defined, wind\_tolerance, reflects the range around the bearing from the facility to the tower location which counts as towards the tower.
-In the base case, we set wind\_tolerance to 10 degrees but we explored setting it to smaller values down to 5 degrees and larger values up to 20 degrees.
-The second hyperparameter we defined, event\_quantile\_effect, is an attempt to deal with the fact that we don't know the travel time of any materials to be deposited via an abnormal release or what if any delays exist between deposition and vegetation response.
-Rather than selecting the maximum value of B3 during a given window or simply the value at the exact time of the event, we set the value of event\_quantile\_effect to correspond with a specific quantile of all B3 values during the event window.
-In the base default case, this was set at 0.9 but we explored values as low as 0.5.
+In the base default case, we set wind\_tolerance to 10 degrees, but we explored setting it to smaller values down to 5 degrees and larger values up to 20 degrees.
+The second hyperparameter we defined, event\_quantile\_effect, handles the unknown travel time of any materials to be deposited via an abnormal release or what, if any, delays exist between deposition and vegetation response.
+Rather than simply selecting the maximum value of $\beta$3 during a given window or the value at the exact time of the event, the value of event\_quantile\_effect is set to correspond with a specific quantile of all $\beta$3 values during the event window.
+In the base default case, this was set at 0.9, but we explored values as low as 0.5.
 All statistical analyses were carried out using the statsmodels Python package (Seabold and Perktold, 2009).
-Our processing scripts are openly available at [doi link], we refer readers to the original providers for data access.
+Our processing scripts are openly available at [doi link], and we refer readers to the original providers for data access.
 
@@
 	\centering
-	\includegraphics[width=9cm]{../figures/__map}
+	\includegraphics[width=16cm]{../figures/__map}
 	\caption{Map of the three closest Fluxnet tower sites to the Fleurus release location.
@@
 \section{Results}
+
+For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51), we found that only the CO2 versus air temperature comparison had any measure of predictability from the model (R2 > 0.05), strong event effect size, and wind direction coming from the release location to towers in ICOS (Table 1).
+Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites.
+When we focus on the period of the abnormal release event in particular rather than the overall period of record, we found that CO2 versus air temperature relationship at each of the three focal towers had similar R2 values and effect sizes (Table 1).
+Figure 2 provides a visual example of the effect size framing in our analysis.
+Note how the apparent interaction effect between time periods around the abnormal release event in Figure 2A (before, during, after) translates to a relatively high effect size (Figure 2C) whereas in Figure 2B at an arbitrary time point, there is no apparent interaction effect.
+This corresponds to a low effect size in Figure 2D.
+
+Despite the similar effect size at the three tower sites during the abnormal release event, they differed substantially in the degree to which they uniquely flagged the known release event.
+At the BE-Lon and BE-Vie sites, our event detection algorithm identified the known release and only a few other time periods as potential abnormal releases (Figure 3).
+Conversely, at the northerly BE-Bra site, the algorithm was not able to identify the known release event distinct from background noise.
+This can be seen in the rate of event detections ($\alpha$) at each site, whereby it was less than 1\% at the BE-Lon and BE-Vie sites, but was greater than 2\% at the BE-Bra site.
+It is notable that this ability to uniquely detect the known release along with a plausible number of additional event detections corresponds with the fraction of the abnormal release period when the wind was pointed from the release location towards the respective tower (using the default base case wind tolerance of 10 deg) which was much lower at the BE-Bra site (5\%) compared to the BE-Lon and BE-Vie sites (34\% Table 1).
+
+To increase our confidence that event detections were not dependent on the specific hyperparameters settings in our base-case setup, we ran an exhaustive cross-validation set of hyperparameter experiments.
+For this effort, we tested every possible combination of values for wind tolerance, number of days, and the event quantile effect.
+Across reasonable values of specific hyperparameters, we found that their magnitudes had little effect on the rate at which events were flagged as potential abnormal releases as even in the most extreme cases, we did not see detection rates exceeding 1\%.
+We did, however, observe that the event detection was related to the window length and event quantile parameters as shown by the slope of the line in Figure 4.
+Although increasing values of the event quantile parameter would hypothetically increase event detection rates, we do not see value in setting this parameter below the median.
+Likewise, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts (e.g.
+Mészáros et al.
+2016).
+
+For the Fukushima case study, our objective was to test the sensitivity of our approach to a broad scale abnormal release and to introduce a control design to verify behavior on a flux tower where we expect no impact from any abnormal releases.
+First, we verified that event detection specificity is present at towers likely to be exposed even across long distances if the abnormal release is large (Figure 5).
+Then, we found a decrease in event detection specificity and an increase in event detection rate moving from the flux tower expected to be most affected (US-Wrc), to a more distant (likely less affected) tower (US-Gle), to a tower likely not exposed at all to the atmospheric plume (OZ-Mul, Figure 5).
 
@@
 	\centering
-	\includegraphics[width=8cm]{../figures/__rolling_fleurus}
+	\includegraphics[width=14cm]{../figures/__rolling_fleurus}
 	\caption{Time series plots of rolling event detection analysis for the CO2 and air temperature (ta) variable pair.
@@
 \section{Discussion}
+Using flux tower records collected closest to the IRE release location, our event detection algorithm was able to identify both a known abnormal release and plausible previously unidentified abnormal releases.
+There was a broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these releases.
+This suggests that given a facility known to have a prior abnormal release, flux tower networks are capable of providing a means of passive monitoring to detect subsequent releases as well as information on the environmental conditions before, during, and after the release.
 
-\section{Data Availability Statement}
+More generally, we show that flux tower records have value for this type of data mining despite the apparent noisiness of the data \citep{fratiniEddyCovarianceFlux2018}.
+We attribute the sensitivity of our approach to the fact that we fit successive interaction models to limited sections of the data record, and we leveraged bivariate relationships to constrain noisiness in individual data records.
+Our approach was aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers.
+This is because although such systematic errors may bias the magnitude of overall flux values, which would be a problem for typical uses of flux towers that deal directly in the units of the measured data, they do not reduce the overall confidence in individual values \citep{langfordEddycovarianceDataLow2015}, which for our purposes might represent an abnormal release signal.
+The ability of our approach to identify a suitable release signature does likely depend on the quality and completeness of the underlying data.
 
-All data and code associated with this manuscript will be made available by the time of publication (or earlier pending the conclusion of US Department of Energy approvals) with a Zenodo DOI at: \href{https://doi.org}{https://doi.org}
+Another contributor to the success of our approach was the fact that Belgium has a relatively dense network of flux towers with long data records.
+As a result, we were able to compare event detections at the closest tower against more distant towers (refer to Figure 1), increasing our confidence that we were seeing a true signal and not spurious noise.
+This level of flux tower density may be found in the United States, Western/Central Europe, Japan, and Australia, which have high densities of flux towers, but not throughout the rest of the world \citep{baldocchi2001fluxnet,pastorello2020fluxnet2015}.
+As a result, this may impose geographical limits on our approach.
+Another attribute that likely aided our investigations was the fact that all three of the flux towers we examined were located in similar agricultural ecosystem types.
+Although investigating whether or not flux tower combinations with differing ecosystems would yield similar results is beyond the scope of the present study, we suspect that the data records among towers in disparate ecosystem types would differ too drastically to be of comparative use.
+For example, \citet{pastorello2020fluxnet2015} show that towers in the crop ecosystem type, of which all three sites used in this study belong, have a relatively narrow distribution of fluxes (i.e., gross primary production, GPP) compared to towers in the grassland or evergreen broadleaf forest types.
+One reason for this narrow spread in crop ecosystems may be the relative homogeneity of flux tower footprints in crop ecosystems, which typically extend to within 1000m of the tower depending on atmospheric conditions (e.g., wind speed and direction, \citep{chuRepresentativenessEddyCovarianceFlux2021}).
+Given that towers in other ecosystem types with a more heterogeneous flux tower footprint have a wider distribution of fluxes (e.g.
+\citep{pastorello2020fluxnet2015}), we suspect that having all the towers in our study located in a homogenous ecosystem type was helpful in reducing the spikiness and spread of the data.
+
+The biggest improvements to our approach would likely come from coupling our data mining procedure with rigorous air pollutant dispersion modeling (e.g., \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
+This would eliminate the need for indirect estimation of material transport via uncertainty analyses and instead use simulation results to analytically determine the likely arrival and duration of material deposition.
+Further improvements could be made by incorporating knowledge about the composition of materials being deposited \citep{international2006iaea, meszarosPredictabilityDispersionFukushimaderived2016} and/or the interactions between material exposure and other environmental stressors \citep{mousseauPlantsLightIonizing2020}, which likely affects the severity and timing of ecosystem responses and by extension the fluxes being measured by a given tower.
+Such information may help disentangle potential interference of atmospheric contaminants with infrared measurement of CO2 from declines in photosynthetic CO2 uptake of contaminated vegetation given our observation that during and after the known abnormal release, the slopes of the relationships between CO2 and air temperature become less negative than before the event, with lower than expected CO2 at low air temperature.
+The identity of possible atmospheric contaminants is unknown.
+We can say, however, that iodine itself is unlikely to interfere with measurements of CO2 given that the absorption spectra of iodine peaks at much shorter wavelengths around 10-7 m \citep{haynes2016crc} compared to that of CO2 which peaks around 1700-2100 10-9 m (LICOR Biosciences Inc., Lincoln, NE).
+
+The fact that we were able to identify a known release (along with potential unidentified releases) in the data record despite not having a detailed radiological or atmospheric transport model is a strength of our approach.
+Because our approach is light-weight and resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration impacts to the plant community within flux tower footprints to ensure human health and industrial safety.
+
+% \section{Data Availability Statement}
+
+% All data and code associated with this manuscript will be made available by the time of publication (or earlier pending the conclusion of US Department of Energy approvals) with a Zenodo DOI at: \href{https://doi.org}{https://doi.org}
 
@@
 
-This are Acknowledgements
+JS built models, analyzed data, and wrote the paper.
+SS, LTD, and AJ contributed to the conception of the manuscript.
+ECT, VAK, LTD, and EC edited the manuscript.
+All authors provided interpretation of results.
+We acknowledge funding for supporting AmeriFlux data portal: U.S.
+Department of Energy Office of Science.
+The data used in this activity has been also funded by CarboEuropeIP (EU-FP6), IMECC (EU-FP6).
+This work was supported by Los Alamos National Laboratory (LDRD-20220062ER).
 

### 2024-08-15 manuscript/manuscript.tex
stash conversion progress

@@
 	\href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Jemma Stachelek$^1$} \\
-	$^1$ Earth and Environmental Sciences\\
+	$^1$ High Performance Computing Division\\
 	Los Alamos National Laboratory\\
@@
 	Vachel A. Kraklow$^2$ \\
-	$^2$ Earth and Environmental Sciences\\
+	$^2$ Earth and Environmental Sciences Division\\
 	Los Alamos National Laboratory\\
@@
 	E. Christi Thompson$^3$ \\
-	$^3$ Earth and Environmental Sciences\\
+	$^3$ Computer, Computational, and Statistical Sciences Division\\
 	Los Alamos National Laboratory\\
@@
 	\textbf{L. Turin Dickman}$^2$ \\
-	$^2$ Earth and Environmental Sciences\\
+	$^2$ Earth and Environmental Sciences Division\\
 	Los Alamos National Laboratory\\
@@
 	\textbf{Emily Casleton}$^3$ \\
-	$^3$ Earth and Environmental Sciences\\
+	$^3$ Computer, Computational, and Statistical Sciences Division\\
 	Los Alamos National Laboratory\\
@@
 	\textbf{Sanna Sevanto}$^2$ \\
-	$^2$ Earth and Environmental Sciences\\
+	$^2$ Earth and Environmental Sciences Division\\
 	Los Alamos National Laboratory\\
@@
 	\textbf{Ann Junghans}$^4$ \\
-	$^4$ Earth and Environmental Sciences\\
+	$^4$ Nuclear and Engineering Nonproliferation Division\\
 	Los Alamos National Laboratory\\
@@
 
-\vspace{-2em}
+\vspace{-4em}
 
@@
 Detecting abnormal releases of radioactive material (hereafter simply “abnormal releases”) is important as a means of ensuring human health and industrial safety.
-The detection of such releases is typically accomplished using active approaches that require site-specific sampling or trace gas measurements (Loyalka 1983) sometimes as a part of international monitoring networks (e.g., Sangiorgi et al.
-2020).
+The detection of such releases is typically accomplished using active approaches that require site-specific sampling or trace gas measurements \citep{loyalkaMechanicsAerosolsNuclear1983} sometimes as a part of international monitoring networks (e.g., \citep{sangiorgiEuropeanRadiologicalData2020}.
 A potential shortcoming of these active monitoring approaches is that they require collection of physical samples (particulates, gasses, and other signal carriers) to retrospectively determine the nature of the release and the events that may have led up to it.
@@
 
-A potential alternative to active monitoring is passive monitoring (e.g., the German Integrated Measuring and Information system [IMIS] and the International Monitoring System [IMS], Bieringer and Schlosser, 2004).
-Although existing wide area monitoring systems are already critical components of abnormal release detection and tracking programs (Medici et al.
-2011), diversification-of-approach may increase the programs’ sensitivity and comprehensiveness.
+A potential alternative to active monitoring is passive monitoring (e.g., the German Integrated Measuring and Information system [IMIS] and the International Monitoring System [IMS], \citep{bieringerMonitoringGroundlevelAir2004}.
+Although existing wide area monitoring systems are already critical components of abnormal release detection and tracking programs \citep{mediciIMSRadionuclideNetwork2001}, diversification-of-approach may increase the programs' sensitivity and comprehensiveness.
 With a greater diversity of methods, monitoring programs may be able to detect a wider array of constituents with differing characteristics.
@@
 Our utilization of flux tower records for event discrimination stands in contrast to the typical use of data from the towers for direct carbon balance accounting.
-Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange (Baldocchi et al.
-2018), our use treats each variable as potentially including an indirect signal (i.e.
+Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange \citep{baldocchiInterannualVariabilityNet2018}, our use treats each variable as potentially including an indirect signal (i.e.
 an anomaly from normal vegetation behavior) of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc).
-In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation in a known way (Nobel 2020).
+In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation in a known way \citep{nobelPhysicochemicalEnvironmentalPlant2020}.
 
@@
 In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008.
-The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq 131I (Carle et al.
-2010).
+The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq 131I \citep{carle2010individual}.
 For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg.
-In the second case study, we further explored potential signals influx tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days eventually dispersing throughout the entire northern hemisphere (Meszaros et al.
-2016).
+In the second case study, we further explored potential signals influx tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days eventually dispersing throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
 We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers.
-While little is known about the direct or indirect effects of radioactive material on vegetation, we expect that exposure may cause declines in photosynthetic CO2 uptake.
-This may be expressed as a change in the slope of the CO2 relationship with other environmental variables (e.g.
+While little is known about the direct or indirect effects of radioactive material on vegetation, we expect that exposure may cause declines in photosynthetic $CO_2$ uptake.
+This may be expressed as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g.
 air temperature, humidity, precipitation, etc.) before, during, and after exposure.
@@
 For the Fukushima case study, we gathered measurements from flux towers located in the Northern Hemisphere (Western United States), which are part of the Ameriflux network (https://ameriflux.lbl.gov/).
-For contrast to the Northern Hemisphere sites where we expect a higher likelihood of impact, we also gathered measurements from flux towers located in the Southern Hemisphere, which are part of the OzFlux network (Cleverly 2011).
+For contrast to the Northern Hemisphere sites where we expect a higher likelihood of impact, we also gathered measurements from flux towers located in the Southern Hemisphere, which are part of the OzFlux network \citep{cleverly2011alice}.
 We selected sites with sufficient long-term data to cover the 2011 date of the disaster and excluded boreal sites with extreme snow cover related seasonality.
@@
 Evaluating an exhaustive list of data transformations was beyond the scope of the present effort.
-After identification of candidate variable pairs, we fit an interaction model of the following form where Y represents a continuous response (e.g., $CO_2$, latent heat, etc.), X1 is an independent variable (e.g., air temperature, humidity, etc.), W2 is a categorical variable representing the “period” of observation (i.e., during, before, or after, the event), $\beta$0 is the intercept, $\beta$1,2 are slope parameters for X1 and W2 respectively, and $\beta$3 is the slope parameter for the interaction between X1 and W2:
-Y = 0+ 1X1 + B2W2+ 3X1W2
+After identification of candidate variable pairs, we fit an interaction model of the following form where Y represents a continuous response (e.g., $CO_2$, latent heat, etc.), $X_1$ is an independent variable (e.g., air temperature, humidity, etc.), $W_2$ is a categorical variable representing the “period” of observation (i.e., during, before, or after, the event), $\beta$0 is the intercept, $\beta$1,2 are slope parameters for $X_1$ and $W_2$ respectively, and $\beta$3 is the slope parameter for the interaction between $X_1$ and $W_2$:
+
+\[ Y = \beta_0 + \beta_1X_1 + \beta_2W_2 + \beta_3X_1W_2 \]
 
@@
 
-For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51), we found that only the CO2 versus air temperature comparison had any measure of predictability from the model (R2 > 0.05), strong event effect size, and wind direction coming from the release location to towers in ICOS (Table 1).
+For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model ($R^2$ > 0.05), strong event effect size, and wind direction coming from the release location to towers in ICOS (Table 1).
 Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites.
-When we focus on the period of the abnormal release event in particular rather than the overall period of record, we found that CO2 versus air temperature relationship at each of the three focal towers had similar R2 values and effect sizes (Table 1).
+When we focus on the period of the abnormal release event in particular rather than the overall period of record, we found that $CO_2$ versus air temperature relationship at each of the three focal towers had similar $R^2$ values and effect sizes (Table 1).
 Figure 2 provides a visual example of the effect size framing in our analysis.
@@
 Likewise, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts (e.g.
-Mészáros et al.
-2016).
+\citep{meszarosPredictabilityDispersionFukushimaderived2016}).
 
@@
 
+\begin{table}[]
+	\centering
+	\caption{Overview of $CO_2$ vs air temperature interaction models at the IRE site.
+$R^2$ is the coefficient of determination of a linear model fit to the two covariates for the period of the abnormal release event (corresponding to “during” in Fig.
+2).
+Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter.
+Wind (t\%) is the average percentage of time that wind direction was from the site of known release towards the selected flux tower during the known release event.
+The remaining columns indicate the base-case hyperparameter settings.}
+	\begin{tabular}{@{}lllllll@{}}
+	\toprule
+	Site & $R^2$ & Effect & Wind(\%) & Days(n) & Wind(tol) & Effect(q) \\ \midrule
+	BE-Vie & 0.35 & 36.2  & 34 & 7 & 10 & 0.9  \\
+	BE-Bra & 0.4 & 57.59  & 5 & 7 & 10 & 0.9 \\
+	BE-Lon & 0.47 & 30.13 & 34 & 7 & 10 & 0.9 \\ \bottomrule
+	\end{tabular}	
+	\label{table:1}
+	\end{table}
+
 \begin{figure}
@@
 	\includegraphics[width=11cm]{../figures/__interaction_belon}
-	\caption{Interaction plots at the BE-Lon site between air temperature (ta) and CO2 (co2) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+	\caption{Interaction plots at the BE-Lon site between air temperature (ta) and $CO_2$ (co2) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
 Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
@@
 	\includegraphics[width=14cm]{../figures/__rolling_fleurus}
-	\caption{Time series plots of rolling event detection analysis for the CO2 and air temperature (ta) variable pair.
+	\caption{Time series plots of rolling event detection analysis for the $CO_2$ and air temperature (ta) variable pair.
 Panel labels indicate the name of the site, the variable pair (in parentheses), and the false positive rate (a).
@@
 Further improvements could be made by incorporating knowledge about the composition of materials being deposited \citep{international2006iaea, meszarosPredictabilityDispersionFukushimaderived2016} and/or the interactions between material exposure and other environmental stressors \citep{mousseauPlantsLightIonizing2020}, which likely affects the severity and timing of ecosystem responses and by extension the fluxes being measured by a given tower.
-Such information may help disentangle potential interference of atmospheric contaminants with infrared measurement of CO2 from declines in photosynthetic CO2 uptake of contaminated vegetation given our observation that during and after the known abnormal release, the slopes of the relationships between CO2 and air temperature become less negative than before the event, with lower than expected CO2 at low air temperature.
+Such information may help disentangle potential interference of atmospheric contaminants with infrared measurement of $CO_2$ from declines in photosynthetic $CO_2$ uptake of contaminated vegetation given our observation that during and after the known abnormal release, the slopes of the relationships between $CO_2$ and air temperature become less negative than before the event, with lower than expected $CO_2$ at low air temperature.
 The identity of possible atmospheric contaminants is unknown.
-We can say, however, that iodine itself is unlikely to interfere with measurements of CO2 given that the absorption spectra of iodine peaks at much shorter wavelengths around 10-7 m \citep{haynes2016crc} compared to that of CO2 which peaks around 1700-2100 10-9 m (LICOR Biosciences Inc., Lincoln, NE).
+We can say, however, that iodine itself is unlikely to interfere with measurements of $CO_2$ given that the absorption spectra of iodine peaks at much shorter wavelengths around 10-7 m \citep{haynes2016crc} compared to that of $CO_2$ which peaks around 1700-2100 10-9 m (LICOR Biosciences Inc., Lincoln, NE).
 

### 2024-08-15 figures/interaction.py
rm title

@@ -137,7 +137,7 @@ gg4.axvline(np.nanmin(pfdist["fdist"]), color="orange")
 ax4.set_xlabel("Effect size")
 ax4.set_title("D.", pad=0, x=0.07, y=0.9)
 #
-plt.suptitle("BE-Lon\nlog(" + var_dep + ") ~ " + var_idep + " * period")
+# plt.suptitle("BE-Lon\nlog(" + var_dep + ") ~ " + var_idep + " * period")
 plt.tight_layout()
 # plt.show()
 plt.savefig("figures/__interaction_belon.pdf", bbox_inches="tight", pad_inches=0.1)

### 2024-08-15 manuscript/manuscript.tex
stash completed conversion

@@
 \usepackage{setspace}
+\usepackage{gensymb}
 
@@
 Detecting abnormal releases of radioactive material (hereafter simply “abnormal releases”) is important as a means of ensuring human health and industrial safety.
-The detection of such releases is typically accomplished using active approaches that require site-specific sampling or trace gas measurements \citep{loyalkaMechanicsAerosolsNuclear1983} sometimes as a part of international monitoring networks (e.g., \citep{sangiorgiEuropeanRadiologicalData2020}.
+The detection of such releases is typically accomplished using active approaches that require site-specific sampling or trace gas measurements \citep{loyalkaMechanicsAerosolsNuclear1983} sometimes as a part of international monitoring networks \citep[e.g.,][]{sangiorgiEuropeanRadiologicalData2020}.
 A potential shortcoming of these active monitoring approaches is that they require collection of physical samples (particulates, gasses, and other signal carriers) to retrospectively determine the nature of the release and the events that may have led up to it.
@@
 
-A potential alternative to active monitoring is passive monitoring (e.g., the German Integrated Measuring and Information system [IMIS] and the International Monitoring System [IMS], \citep{bieringerMonitoringGroundlevelAir2004}.
+A potential alternative to active monitoring is passive monitoring \citep[e.g., the German Integrated Measuring and Information system and the International Monitoring System,][]{bieringerMonitoringGroundlevelAir2004}.
 Although existing wide area monitoring systems are already critical components of abnormal release detection and tracking programs \citep{mediciIMSRadionuclideNetwork2001}, diversification-of-approach may increase the programs' sensitivity and comprehensiveness.
@@
 Although increasing values of the event quantile parameter would hypothetically increase event detection rates, we do not see value in setting this parameter below the median.
-Likewise, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts (e.g.
-\citep{meszarosPredictabilityDispersionFukushimaderived2016}).
+Likewise, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.
 
@@
 	\includegraphics[width=11cm]{../figures/__interaction_belon}
-	\caption{Interaction plots at the BE-Lon site between air temperature (ta) and $CO_2$ (co2) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+	\caption{Interaction plots at the BE-Lon site between air temperature ($C^{\circ}$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
 Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
@@
 	\centering
-	\includegraphics[width=14cm]{../figures/__rolling_fleurus}
-	\caption{Time series plots of rolling event detection analysis for the $CO_2$ and air temperature (ta) variable pair.
-Panel labels indicate the name of the site, the variable pair (in parentheses), and the false positive rate (a).
-Event detection lines (orange), interaction effect size (blue), and event effect size (green) correspond to the left vertical axis.
-Fraction that the wind direction was towards the particular tower (black) corresponds to the right vertical axis.}
+	\includegraphics[width=16cm]{../figures/__rolling_fleurus}
+	\caption{Time series plots (2004-2013) of rolling event detection analysis at the IRE site for the $CO_2$ and air temperature ($C^{\circ}$) variable pair.
+Event detection lines (orange), interaction effect size (blue), and event effect size (dashed black line) are shown in the top panels.
+The fraction of time that the wind direction was towards the particular tower (solid black) are shown in the corresponding bottom panels.}
 	\label{fig:rolling}
@@
 	\centering
-	\includegraphics[width=12cm]{../figures/__hyperparameter_experiment}
-	\caption{Sensitivity of false event detection to different hyperparameter values.}
+	\includegraphics[width=16cm]{../figures/__hyperparameter_experiment}
+	\caption{Sensitivity of event detection at the IRE site to different hyperparameter values (points) alongside potential relationship (solid line) and confidence interval (shaded region).}
 	\label{fig:hyperparameter}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=16cm]{../figures/__rolling_fukushima}
+	\caption{Time series plots (2010-2013) of rolling event detection analysis for the Fukushima disaster case study.
+The depicted variable pair is latent heat and relative humidity.
+Event detection lines (orange), interaction effect size (blue), and event effect size (dashed black line) are shown in the top panels.
+The fraction of time that the wind direction was towards the particular tower (solid black) are shown in the corresponding bottom panels.}
+	\label{fig:fukushima}
 \end{figure}
@@
 For example, \citet{pastorello2020fluxnet2015} show that towers in the crop ecosystem type, of which all three sites used in this study belong, have a relatively narrow distribution of fluxes (i.e., gross primary production, GPP) compared to towers in the grassland or evergreen broadleaf forest types.
-One reason for this narrow spread in crop ecosystems may be the relative homogeneity of flux tower footprints in crop ecosystems, which typically extend to within 1000m of the tower depending on atmospheric conditions (e.g., wind speed and direction, \citep{chuRepresentativenessEddyCovarianceFlux2021}).
-Given that towers in other ecosystem types with a more heterogeneous flux tower footprint have a wider distribution of fluxes (e.g.
-\citep{pastorello2020fluxnet2015}), we suspect that having all the towers in our study located in a homogenous ecosystem type was helpful in reducing the spikiness and spread of the data.
+One reason for this narrow spread in crop ecosystems may be the relative homogeneity of flux tower footprints in crop ecosystems, which typically extend to within 1000m of the tower depending on atmospheric conditions \citep[e.g., wind speed and direction,][]{chuRepresentativenessEddyCovarianceFlux2021}.
+Given that towers in other ecosystem types with a more heterogeneous flux tower footprint have a wider distribution of fluxes \citep[e.g.][]{pastorello2020fluxnet2015}, we suspect that having all the towers in our study located in a homogenous ecosystem type was helpful in reducing the spikiness and spread of the data.
 
-The biggest improvements to our approach would likely come from coupling our data mining procedure with rigorous air pollutant dispersion modeling (e.g., \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
+The biggest improvements to our approach would likely come from coupling our data mining procedure with rigorous air pollutant dispersion modeling \cite[e.g.,][]{meszarosPredictabilityDispersionFukushimaderived2016}.
 This would eliminate the need for indirect estimation of material transport via uncertainty analyses and instead use simulation results to analytically determine the likely arrival and duration of material deposition.

### 2024-08-15 manuscript/manuscript.tex
reference supplementary table

@@
 
-For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model ($R^2$ > 0.05), strong event effect size, and wind direction coming from the release location to towers in ICOS (Table 1).
+For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51, Table S1), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model ($R^2$ > 0.05), strong event effect size, and wind direction coming from the release location to towers in ICOS (Table 1).
 Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites.

### 2024-08-19 manuscript/manuscript.tex
text edits

--- 2024-08-15 manuscript/manuscript.tex
@@
 As an alternative first step towards passive monitoring, we developed an approach using eddy flux tower data records to identify signals from a known abnormal release and quantify the extent to which that signal also occurs at other times in the data record.
-Through two case-studies, one of which targeted the Fukushima nuclear disaster and the other targeting a radioisotope production facility, we tested our approach and identified several potential heretofore unidentified abnormal releases that were consistent with atmospheric circulation patterns and/or wind direction from known release sites.
-Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
+Through two case-studies, one of which targeted the Fukushima nuclear disaster and the other targeting an abnormal release event at a radioisotope production facility in Fleurus Belgium, we tested our approach and identified several potential heretofore unidentified abnormal releases that were consistent with atmospheric circulation patterns and/or wind direction from known release sites.Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
 
@@
 For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg.
-In the second case study, we further explored potential signals influx tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days eventually dispersing throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
+In the second case study, we further explored potential signals influx tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
 We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers.
-While little is known about the direct or indirect effects of radioactive material on vegetation, we expect that exposure may cause declines in photosynthetic $CO_2$ uptake.
+While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, we expect that exposure may cause declines in photosynthetic $CO_2$ uptake.
 This may be expressed as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g.
@@
 For the IRE case study, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/).
-Only three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1).
+Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1).
 Each of the three flux tower sites are located within a crop ecosystem type and had almost 10 years of data available from 2004 through 2013 recorded at 30 min intervals.
@@
 
-For the Fukushima case study, we gathered measurements from flux towers located in the Northern Hemisphere (Western United States), which are part of the Ameriflux network (https://ameriflux.lbl.gov/).
+For the Fukushima case study, due to the direction of the plume, we gathered measurements from flux towers located in the Northern Hemisphere (Western United States), which are part of the Ameriflux network (https://ameriflux.lbl.gov/).
 For contrast to the Northern Hemisphere sites where we expect a higher likelihood of impact, we also gathered measurements from flux towers located in the Southern Hemisphere, which are part of the OzFlux network \citep{cleverly2011alice}.
@@
 We compared the effect size of the interaction term ($\beta$3) during the event-encompassing window against all other windows in a “rolling” analysis.
-Therefore, the overall analysis compares the interaction term of the window of the known event (which occurred on August 23, 2008) against a window centered on every other observation.
+Therefore, the overall analysis compares the interaction term of the window of the known event against a window centered on every other observation.
 Specific windows that had an effect size at least as large as that of the known event and with wind conditions defined as “towards” the flux tower were flagged as “event detections”.
-The frequency of these event detections are reported as “event detections”.
 
@@
 
-For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model ($R^2$ > 0.05), strong event effect size, and wind direction coming from the release location to towers in ICOS (Table 1).
+For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51, Table S1), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model ($R^2$ > 0.05) and a strong event effect size (Table 1).
 Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites.
@@
 Another attribute that likely aided our investigations was the fact that all three of the flux towers we examined were located in similar agricultural ecosystem types.
-Although investigating whether or not flux tower combinations with differing ecosystems would yield similar results is beyond the scope of the present study, we suspect that the data records among towers in disparate ecosystem types would differ too drastically to be of comparative use.
+Although investigating whether or not flux tower combinations with differing ecosystems would yield similar results is beyond the scope of the present study, we suspect that the data records among towers in disparate ecosystem types could differ too drastically to be of comparative use.
 For example, \citet{pastorello2020fluxnet2015} show that towers in the crop ecosystem type, of which all three sites used in this study belong, have a relatively narrow distribution of fluxes (i.e., gross primary production, GPP) compared to towers in the grassland or evergreen broadleaf forest types.

### 2024-08-20 manuscript/manuscript.tex
stash text edits

@@
 This signal is then used to investigate whether additional potential abnormal releases exist at other times in the data record.
-We focused on eddy flux towers, as they are a potential source of abnormal release signatures, are globally distributed, have been in near continuous operation for several decades, and can be relatively cheaply and easily deployed in locations of interest.
+We focused on eddy flux towers, as they have to potential to provide abnormal release signatures, are globally distributed, have been in near continuous operation for several decades, and can be relatively cheaply and easily deployed in locations of interest.
 Therefore, they constitute a rich target for data mining approaches aimed at identifying signatures resulting from direct interference of radioactive material with target measurements and/or the indirect signal of vegetation responses to the deposition of radioactive material.

### 2024-08-21 figures/maps.R
linting

@@ -103,23 +103,28 @@ ggsave("figures/__map_mul.pdf", gg)
 
 # --- fukushima
 # https://r-spatial.org/r/2018/10/25/ggplot2-sf.html
-crsrobin <- "+proj=robin +lon_0=-180 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
+crsrobin <- paste0("+proj=robin +lon_0=-180 +x_0=0 +y_0=0 +ellps=WGS84",
+  " +datum=WGS84 +units=m +no_defs")
 
 dt_fuku_df <- data.frame(
   site_code = c("Fukushima", "US-Wrc", "US-Gle", "OZ-Mul"),
   lat = c(37.423056, 45.8205, 41.3665, -22.2828),
   lon = c(141.033056, -121.9519, -106.2399, 133.2493))
-dt_fuku <- st_transform(st_as_sf(dt_fuku_df, coords = c("lon", "lat"), crs = 4326), crs = crsrobin)
+dt_fuku <- st_transform(
+  st_as_sf(dt_fuku_df, coords = c("lon", "lat"), crs = 4326), crs = crsrobin)
 dt_fuku_df <- cbind(dt_fuku_df, st_coordinates(dt_fuku))
 
 library(rnaturalearth)
 world <- ne_countries(scale = "medium", returnclass = "sf")
-gg_overview <- ggplot(data = st_transform(st_break_antimeridian(world, lon_0 = 180), crs = crsrobin)) +
+gg_overview <- ggplot(
+  data = st_transform(st_break_antimeridian(world, lon_0 = 180), crs = crsrobin)) +
   geom_sf() +
   geom_sf(data = dt_fuku) +
   geom_text(data = dt_fuku_df, aes(x = X, y = Y * 1.06, label = site_code),
     color = "darkblue", fontface = "bold", check_overlap = FALSE) +
-  coord_sf(xlim = c(-4320952 * 1.3, 6376465  * 1.2), ylim = c(-2383193 * 1.3, 4891261 * 1.3), expand = FALSE) +
+  coord_sf(
+    xlim = c(-4320952 * 1.3, 6376465  * 1.2),
+    ylim = c(-2383193 * 1.3, 4891261 * 1.3), expand = FALSE) +
   theme_bw() +
   theme(axis.title = element_blank())
 gg <- cowplot::plot_grid(gg_overview, gg_us_wrc, gg_oz_mul, gg_us_gle)

### 2024-08-21 figures/maps.R
map styling

@@ -49,7 +49,8 @@ gg_euro_bevie <- get_gg_sub(dt_sub, label_color = "white")
 dt_sub <- dplyr::filter(m1_data, site_code == "BE-Bra")
 gg_euro_bebra <- get_gg_sub(dt_sub, label_color = "white")
 
-gg <- cowplot::plot_grid(gg_overview, gg_euro_bebra, gg_euro_belon, gg_euro_bevie)
+gg <- cowplot::plot_grid(gg_overview, gg_euro_bebra, gg_euro_belon, gg_euro_bevie) +
+  cowplot::panel_border(color = "black", size = 1.5)
 ggsave("figures/__map_fleurus.pdf", gg)
 
 # --- jp-fhk
@@ -116,16 +117,34 @@ dt_fuku_df <- cbind(dt_fuku_df, st_coordinates(dt_fuku))
 
 library(rnaturalearth)
 world <- ne_countries(scale = "medium", returnclass = "sf")
+xmin <- -4320952
+xmax <- 6376465
+ymin <- -2383193 * 1.12
+ymax <- 4891261 * 1.12
+arrow_y <- ((xmin + xmax) / 2) + abs(ymin - ymax) * 0.32
+arrow_x_end <- ((xmin + xmax) / 2) + abs(xmin - xmax) * 0.16
 gg_overview <- ggplot(
   data = st_transform(st_break_antimeridian(world, lon_0 = 180), crs = crsrobin)) +
   geom_sf() +
-  geom_sf(data = dt_fuku) +
+  geom_sf(data = dt_fuku, color = "red") +
   geom_text(data = dt_fuku_df, aes(x = X, y = Y * 1.06, label = site_code),
-    color = "darkblue", fontface = "bold", check_overlap = FALSE) +
+    check_overlap = FALSE) +
   coord_sf(
-    xlim = c(-4320952 * 1.3, 6376465  * 1.2),
-    ylim = c(-2383193 * 1.3, 4891261 * 1.3), expand = FALSE) +
-  theme_bw() +
-  theme(axis.title = element_blank())
-gg <- cowplot::plot_grid(gg_overview, gg_us_wrc, gg_oz_mul, gg_us_gle)
+    xlim = c(xmin * 1.2, xmax  * 1.1),
+    ylim = c(ymin * 1.3, ymax * 1.3), expand = TRUE) +
+  annotate("text", x = (2 + arrow_x_end) / 2, y = arrow_y - abs(ymin - ymax) * 0.1,
[…]

### 2024-08-21 manuscript/manuscript.tex
text edits

@@
 Although existing wide area monitoring systems are already critical components of abnormal release detection and tracking programs \citep{mediciIMSRadionuclideNetwork2001}, diversification-of-approach may increase the programs' sensitivity and comprehensiveness.
-With a greater diversity of methods, monitoring programs may be able to detect a wider array of constituents with differing characteristics.
+With a greater diversity of methods, monitoring programs may be able to detect a wider array of constituents with differing characteristics or expand their geographic extent without the need to deploy new instrumentation.
 In the present study, we developed a novel passive monitoring approach to supplement existing programs that involves the use of eddy covariance data to capture a signal from known abnormal releases.
@@
 Such information may help disentangle potential interference of atmospheric contaminants with infrared measurement of $CO_2$ from declines in photosynthetic $CO_2$ uptake of contaminated vegetation given our observation that during and after the known abnormal release, the slopes of the relationships between $CO_2$ and air temperature become less negative than before the event, with lower than expected $CO_2$ at low air temperature.
-The identity of possible atmospheric contaminants is unknown.
-We can say, however, that iodine itself is unlikely to interfere with measurements of $CO_2$ given that the absorption spectra of iodine peaks at much shorter wavelengths around 10-7 m \citep{haynes2016crc} compared to that of $CO_2$ which peaks around 1700-2100 10-9 m (LICOR Biosciences Inc., Lincoln, NE).
+Although the chemical makeup of possible atmospheric contaminants is unknown, interference by iodine itself is unlikely to affect measurements of $CO_2$ given that the absorption spectra of iodine peaks at much shorter wavelengths around 10-7 m \citep{haynes2016crc} compared to that of $CO_2$ which peaks around 1700-2100 10-9 m (LICOR Biosciences Inc., Lincoln, NE).
 
-The fact that we were able to identify a known release (along with potential unidentified releases) in the data record despite not having a detailed radiological or atmospheric transport model is a strength of our approach.
+The fact that we were able to identify a known release (along with potential unidentified abnormal events) in the data record despite not having a detailed radiological or atmospheric transport model is a strength of our approach.
+Furthermore, our approach is not limited to a radiological context, any abnormal event that affects the plant community and is reflected in flux tower data is a potential target.
 Because our approach is light-weight and resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration impacts to the plant community within flux tower footprints to ensure human health and industrial safety.

### 2024-08-22 manuscript/manuscript.tex
stash text edits

@@
 As an alternative first step towards passive monitoring, we developed an approach using eddy flux tower data records to identify signals from a known abnormal release and quantify the extent to which that signal also occurs at other times in the data record.
-Through two case-studies, one of which targeted the Fukushima nuclear disaster and the other targeting an abnormal release event at a radioisotope production facility in Fleurus Belgium, we tested our approach and identified several potential heretofore unidentified abnormal releases that were consistent with atmospheric circulation patterns and/or wind direction from known release sites.Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
+Through two case-studies, one of which targeted the Fukushima nuclear disaster and the other targeting an abnormal release event at a radioisotope production facility in Fleurus Belgium, we tested our approach and identified several potential heretofore unidentified abnormal events that were consistent with atmospheric circulation patterns and/or wind direction from known release sites.Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
 
@@
 \section{Discussion}
-Using flux tower records collected closest to the IRE release location, our event detection algorithm was able to identify both a known abnormal release and plausible previously unidentified abnormal releases.
-There was a broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these releases.
-This suggests that given a facility known to have a prior abnormal release, flux tower networks are capable of providing a means of passive monitoring to detect subsequent releases as well as information on the environmental conditions before, during, and after the release.
+Using flux tower records collected closest to the IRE release location, our event detection algorithm was able to identify both a known abnormal release and plausible previously unidentified abnormal events.
+There was a broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these events.
+This suggests that, at the very least, given a facility known to have a prior abnormal event, flux tower networks are capable of providing a means of passive monitoring to detect subsequent events as well as information on the environmental conditions before, during, and after the event.
+Furthermore, our approach has promise for situations when there is not a prior known abnormal release.
+In these cases, it may be possible to continuously mine flux tower records (as they become available) for events that exceed some threshold identified in a more comprehensive cross-site, cross-event benchmarking study.
 

### 2024-08-22 manuscript/manuscript.tex
stash text edits

@@
 	\includegraphics[width=16cm]{../figures/__map}
-	\caption{Map of the three closest Fluxnet tower sites to the Fleurus release location.
-Note that BE-Lon and BE-Vie are of a similar easterly direction to the site whereas BE-Bra is of an entirely different northerly direction.}
+	\caption{Map of flux tower locations for the IRE and Fukushima case studies (left and right respectively).
+Note that BE-Lon and BE-Vie are of a similar easterly direction to the IRE site whereas BE-Bra is in a northerly direction.
+The distance between IRE and the flux tower sites are approximately 19, 95, and 105 km for the BE-Lon, BE-Bra, and BE-Vie respectively.
+Each of the three IRE flux tower sites are located within a crop ecosystem type.
+Note that the US-Wrc site would be expected to have first contact with the Fukushima atmospheric plume followed by US-Gle.
+The OZ-Mul site would not be expected to have been affected by the plume.}
 	\label{fig:study_site}

### 2024-08-28 manuscript/manuscript.tex
text edits

@@
 As an alternative first step towards passive monitoring, we developed an approach using eddy flux tower data records to identify signals from a known abnormal release and quantify the extent to which that signal also occurs at other times in the data record.
-Through two case-studies, one of which targeted the Fukushima nuclear disaster and the other targeting an abnormal release event at a radioisotope production facility in Fleurus Belgium, we tested our approach and identified several potential heretofore unidentified abnormal events that were consistent with atmospheric circulation patterns and/or wind direction from known release sites.Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
+Through two case-studies, one of which targeted the Fukushima nuclear disaster and the other targeting an abnormal release event at a radioisotope production facility in Fleurus Belgium, we tested our approach and identified several potential heretofore unidentified abnormal events that were consistent with atmospheric circulation patterns and/or wind direction from known release sites.
+Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
 
@@
 For the Fukushima case study, due to the direction of the plume, we gathered measurements from flux towers located in the Northern Hemisphere (Western United States), which are part of the Ameriflux network (https://ameriflux.lbl.gov/).
-For contrast to the Northern Hemisphere sites where we expect a higher likelihood of impact, we also gathered measurements from flux towers located in the Southern Hemisphere, which are part of the OzFlux network \citep{cleverly2011alice}.
+To increase confidence that we are seeing a real effect, we contrasted the results from Northern Hemisphere sites, where we expect a higher likelihood of impact, to measurements from flux towers located in the Southern Hemisphere, which are part of the OzFlux network \citep{cleverly2011alice}.
 We selected sites with sufficient long-term data to cover the 2011 date of the disaster and excluded boreal sites with extreme snow cover related seasonality.

### 2024-08-28 figures/maps.R
add map scalebars

@@ -1,4 +1,6 @@
+options(debug = TRUE)
 library(ggmap) # install_github("stadiamaps/ggmap")
+library(ggspatial)
 source("scripts/99_utils.R")
 
 generate_overview <- function(sites, site_codes, buffer_x = 0.5, buffer_y = 0.2,
@@ -21,6 +23,17 @@ generate_overview <- function(sites, site_codes, buffer_x = 0.5, buffer_y = 0.2,
 
   gg_map <- get_map(location = bbox,
     maptype  = "stamen_toner_background")
+
+  # https://stackoverflow.com/a/20580465
+  bb <- attr(gg_map, "bb")
+  sbar <- data.frame(lon.start = c(bb$ll.lon + 0.1 * (bb$ur.lon - bb$ll.lon)),
+    lon.end = c(bb$ll.lon + 0.25 * (bb$ur.lon - bb$ll.lon)),
+    lat.start = c(bb$ll.lat + 0.1 * (bb$ur.lat - bb$ll.lat)),
+    lat.end = c(bb$ll.lat + 0.1 * (bb$ur.lat - bb$ll.lat)))
+  sbar$distance <- dist_haversine(long = c(sbar$lon.start, sbar$lon.end),
+    lat = c(sbar$lat.start, sbar$lat.end))
+  ptspermm <- 2.83464567  # need this because geom_text uses mm, and themes use pts. Urgh.
+
   gg_overview <- ggmap(gg_map) +
     geom_point(
       data = m1_data, aes(x = X, y = Y),
@@ -32,7 +45,26 @@ generate_overview <- function(sites, site_codes, buffer_x = 0.5, buffer_y = 0.2,
       hjust = nudge_x,
       size = 6,
       color = label_color
-    )
+    ) +
+    geom_segment(data = sbar,
+      aes(x = lon.start,
+        xend = lon.end,
+        y = lat.start,
+        yend = lat.end),
+      arrow = arrow(angle = 90, length = unit(0.1, "cm"),
+        ends = "both", type = "open"),
+      color = label_color) +
+    geom_text(data = sbar,
+      aes(x = (lon.start + lon.end) / 2,
+        y = lat.start + 0.025 * (bb$ur.lat - bb$ll.lat),
+        label = paste(format(distance,
+          digits = 4,
+          nsmall = 2),
+        "km")),
+      hjust = 0.5,
+      vjust = 0,
+      size = 8 / ptspermm,
[…]

### 2024-08-28 scripts/99_utils.R

@@ -48,6 +48,17 @@ get_gg_sub <- function(dt_sub, buffer_x = 0.02, buffer_y = 0.02, label_color = "
   gg_map <- get_googlemap(center = c(lon = dt_sub$X, lat = dt_sub$Y),
     maptype  = "satellite",
     zoom     = 16)
+
+  # https://stackoverflow.com/a/20580465
+  bb <- attr(gg_map, "bb")
+  sbar <- data.frame(lon.start = c(bb$ll.lon + 0.1 * (bb$ur.lon - bb$ll.lon)),
+    lon.end = c(bb$ll.lon + 0.25 * (bb$ur.lon - bb$ll.lon)),
+    lat.start = c(bb$ll.lat + 0.1 * (bb$ur.lat - bb$ll.lat)),
+    lat.end = c(bb$ll.lat + 0.1 * (bb$ur.lat - bb$ll.lat)))
+  sbar$distance <- dist_haversine(long = c(sbar$lon.start, sbar$lon.end),
+    lat = c(sbar$lat.start, sbar$lat.end))
+  ptspermm <- 2.83464567  # need this because geom_text uses mm, and themes use pts. Urgh.
+
   res <- ggmap(gg_map) +
     geom_point(aes(x = X, y = Y),
       data = dt_sub,
@@ -58,6 +69,25 @@ get_gg_sub <- function(dt_sub, buffer_x = 0.02, buffer_y = 0.02, label_color = "
       vjust = vjust,
       hjust = hjust,
       size = 6,
+      color = label_color) +
+    geom_segment(data = sbar,
+      aes(x = lon.start,
+        xend = lon.end,
+        y = lat.start,
+        yend = lat.end),
+      arrow = arrow(angle = 90, length = unit(0.1, "cm"),
+        ends = "both", type = "open"),
+      color = label_color) +
+    geom_text(data = sbar,
+      aes(x = (lon.start + lon.end) / 2,
+        y = lat.start + 0.025 * (bb$ur.lat - bb$ll.lat),
+        label = paste(format(distance,
+          digits = 4,
+          nsmall = 2),
+        "km")),
+      hjust = 0.5,
+      vjust = 0,
+      size = 8 / ptspermm,
       color = label_color)
   res
 }
@@ -94,4 +124,19 @@ get_sites <- function(lat, lon, site_tag, distance_threshold = 100) {
[…]

### 2024-08-28 manuscript/manuscript.tex
text edit

@@
 
-The fact that we were able to identify a known release (along with potential unidentified abnormal events) in the data record despite not having a detailed radiological or atmospheric transport model is a strength of our approach.
+The fact that we were able to identify known releases (along with potential unidentified abnormal events) in the data record from both a large event (i.e.
+Fukushima) as well as a smaller event despite not having a detailed radiological or atmospheric transport model is a strength of our approach.
 Furthermore, our approach is not limited to a radiological context, any abnormal event that affects the plant community and is reflected in flux tower data is a potential target.

### 2024-08-29 manuscript/manuscript.tex
text edits

@@
 This signal is then used to investigate whether additional potential abnormal releases exist at other times in the data record.
-We focused on eddy flux towers, as they have to potential to provide abnormal release signatures, are globally distributed, have been in near continuous operation for several decades, and can be relatively cheaply and easily deployed in locations of interest.
+We focused on eddy flux towers, as they have the potential to provide abnormal release signatures, are globally distributed, have been in near continuous operation for several decades, and can be relatively cheaply and easily deployed in locations of interest.
 Therefore, they constitute a rich target for data mining approaches aimed at identifying signatures resulting from direct interference of radioactive material with target measurements and/or the indirect signal of vegetation responses to the deposition of radioactive material.
@@
 For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg.
-In the second case study, we further explored potential signals influx tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
+In the second case study, we further explored potential signals in flux tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
 We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers.
@@
 For the IRE case study, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/).
-Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1).
+Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1A).
 Each of the three flux tower sites are located within a crop ecosystem type and had almost 10 years of data available from 2004 through 2013 recorded at 30 min intervals.
@@
 Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter.
-Wind (t\%) is the average percentage of time that wind direction was from the site of known release towards the selected flux tower during the known release event.
+Wind (\%) is the average percentage of time that wind direction was from the site of known release towards the selected flux tower during the known release event.
 The remaining columns indicate the base-case hyperparameter settings.}
@@
 \section{Discussion}
-Using flux tower records collected closest to the IRE release location, our event detection algorithm was able to identify both a known abnormal release and plausible previously unidentified abnormal events.
-There was a broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these events.
-This suggests that, at the very least, given a facility known to have a prior abnormal event, flux tower networks are capable of providing a means of passive monitoring to detect subsequent events as well as information on the environmental conditions before, during, and after the event.
+Using flux tower records collected closest to and in the prevailing wind direction to the IRE and the Fukushima release locations, our event detection algorithm was able to identify both known abnormal releases and plausible previously unidentified abnormal events.
+There was broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these releases, and our approach was supported by the loss of signal at flux towers outside of the prevailing wind direction.
+This suggests that, at the very least, given a facility known to have a prior abnormal release, flux tower networks are capable of providing a means of passive monitoring to detect subsequent events as well as information on the environmental conditions before, during, and after the event.
 Furthermore, our approach has promise for situations when there is not a prior known abnormal release.
@@
 
-Another contributor to the success of our approach was the fact that Belgium has a relatively dense network of flux towers with long data records.
-As a result, we were able to compare event detections at the closest tower against more distant towers (refer to Figure 1), increasing our confidence that we were seeing a true signal and not spurious noise.
+Another contributor to the success of our approach was our ability to compare event detections at  close towers in prevailing wind directions against more distant towers (refer to Figure 1), increasing our confidence that we were seeing a true signal and not spurious noise.
+The relatively dense networks of flux towers with long data records leveraged in both locations also aided these case studies.
 This level of flux tower density may be found in the United States, Western/Central Europe, Japan, and Australia, which have high densities of flux towers, but not throughout the rest of the world \citep{baldocchi2001fluxnet,pastorello2020fluxnet2015}.
-As a result, this may impose geographical limits on our approach.
-Another attribute that likely aided our investigations was the fact that all three of the flux towers we examined were located in similar agricultural ecosystem types.
+As a result, this may impose geographical limits on our approach, although installation and maintenance of new flux towers in locations of interest may be cost effective compared to other monitoring approaches.
+Another attribute that likely aided our investigations was the fact that all three of the flux towers we examined were located in similar agricultural and forested ecosystem types for the IRE and Fukushima case studies respectively.
 Although investigating whether or not flux tower combinations with differing ecosystems would yield similar results is beyond the scope of the present study, we suspect that the data records among towers in disparate ecosystem types could differ too drastically to be of comparative use.
@@
 One reason for this narrow spread in crop ecosystems may be the relative homogeneity of flux tower footprints in crop ecosystems, which typically extend to within 1000m of the tower depending on atmospheric conditions \citep[e.g., wind speed and direction,][]{chuRepresentativenessEddyCovarianceFlux2021}.
-Given that towers in other ecosystem types with a more heterogeneous flux tower footprint have a wider distribution of fluxes \citep[e.g.][]{pastorello2020fluxnet2015}, we suspect that having all the towers in our study located in a homogenous ecosystem type was helpful in reducing the spikiness and spread of the data.
+Given that towers in other ecosystem types with a more heterogeneous flux tower footprint have a wider distribution of fluxes \citep[e.g.][]{pastorello2020fluxnet2015}, we suspect that having all the towers in the IRE case study located in a homogenous ecosystem type was helpful in reducing the spikiness and spread of the data.
 
@@
 
-The fact that we were able to identify known releases (along with potential unidentified abnormal events) in the data record from both a large event (i.e.
+The fact that we were able to identify known releases (along with potential unidentified abnormal events) in the data records from both a large event (i.e.
 Fukushima) as well as a smaller event despite not having a detailed radiological or atmospheric transport model is a strength of our approach.

### 2024-08-29 figures/maps.R
figure cleanup

@@ -69,7 +69,7 @@ generate_overview <- function(sites, site_codes, buffer_x = 0.5, buffer_y = 0.2,
 }
 
 # --- ire
-sites <- get_sites(50.45055, 4.5350415, "fleurus")
+sites <- get_sites(50.45055, 4.5350415, "IRE")
 overview <- generate_overview(sites, c("BE-Lon", "BE-Vie", "BE-Bra", "IRE"))
 gg_overview <- overview$gg_overview
 # ggsave("test.pdf", gg_overview)

### 2024-08-29 scripts/01_fit_rolling.py

@@ -280,7 +280,6 @@ def fit_rolling(
         ax2 = axs[1]
 
         g = sns.lineplot(data=g_data, x="timestamp", y="F", ax=ax1)
-        g.axvline(pd.to_datetime(date_event), color="yellow")
         g.axhline(event_effect, color="black", ls="--")
         g.set_ylim(0, panel_ylim)
         # ax1.xaxis.set_major_formatter(mdates.DateFormatter("%Y"))
@@ -358,6 +357,10 @@ def fit_rolling(
         # g_data[
         #     (g_data["p"] > abs(np.log(p_event))).values and (g_data["test"] > 0.7).values
         # ].shape
+
+        g.axvline(pd.to_datetime(date_event), color="yellow")
+        g2.axvline(pd.to_datetime(date_event), color="yellow")
+
         if noyticklabels:
             ax2.set_yticklabels([])
             ax2.yaxis.label.set_color("white")
@@ -378,7 +381,7 @@ def fit_rolling(
             x=0.25,
         )
         ax1.set_xlabel("")
-        ax2.set_xlabel("")
+        ax2.set_xlabel("Year")
         # plt.show()
         print(path_fig)
         plt.subplots_adjust(hspace=0.1)

### 2024-08-29 manuscript/manuscript.tex
no plans to release data
late breaking text edits

@@
 
-We specifically focused on short-term responses of vegetation (on the order of days to weeks) to potential deposition of radioactive material rather than chronic long-term signals (on the order of years to decades).
+We specifically focused on short-term anomalies in the vegetation response data (on the order of days to weeks) due to either potential deposition of radioactive material or direct sensor effects rather than chronic long-term signals on the order of years to decades.
 In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008.
@@
 In the second case study, we further explored potential signals in flux tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
-We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers.
+We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other eddy flux towers.
 While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, we expect that exposure may cause declines in photosynthetic $CO_2$ uptake.
@@
 While BE-Lon had a near-continuous data record during this period, BE-Bra had a long period of missing data in 2005, and both BE-Bra and BE-Vie had a long period of missing data in 2009.
-The distance between IRE and the flux tower sites are approximately 19 km, 95 km, and 105 km for the BE-Lon, BE-Bra, and BE-Vie respectively.
+The distance between IRE and the flux tower sites are approximately 19 km, 95 km, and 105 km for the BE-Lon, BE-Bra, and BE-Vie, respectively.
 
@@
 
-This interaction model is fit for a window of time encompassing the known release event and for every other possible window in the period-of-record.
+This interaction model was fit for a window of time encompassing the known release event and for every other possible window in the period-of-record.
 The overall length of the window was treated as an adjustable hyperparameter since the length of time for an abnormal material release to reach the footprint of the flux tower may be event-dependent and/or unknown.
-In addition, the length of time for deposited material to induce a possible vegetation response (or at least a flux tower anomaly) is an unknown quantity.
+In addition, the length of time for deposited material to induce a possible vegetation response (or at least a flux data anomaly) is an unknown quantity.
 Thus, in the default base case, we set the window length (n\_days) as 7 days, but we ran hyperparameter experiments to determine the result of setting it to longer periods of 10 and 14 days.
@@
 
-In order to flag event detections, we defined several adjustable hyperparameters to deal with uncertainty surrounding atmospheric transport and the timing of any possible responses to material deposition.
+In order to flag event detections, we defined several adjustable hyperparameters to account for uncertainty surrounding atmospheric transport and the timing of any possible responses to material deposition.
 The first hyperparameter we defined, wind\_tolerance, reflects the range around the bearing from the facility to the tower location which counts as towards the tower.
@@
 This corresponds to a low effect size in Figure 2D.
+Note also how during and after the known abnormal release, the slopes of the relationships between CO2 and air temperature become less negative than before the event, with lower than expected CO2 at low air temperature (Figure 2A).
 
@@
 Although increasing values of the event quantile parameter would hypothetically increase event detection rates, we do not see value in setting this parameter below the median.
-Likewise, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.
+Similarly, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.
 
@@
 
-% All data and code associated with this manuscript will be made available by the time of publication (or earlier pending the conclusion of US Department of Energy approvals) with a Zenodo DOI at: \href{https://doi.org}{https://doi.org}
+% All code associated with this manuscript will be made available by the time of publication (or earlier pending the conclusion of US Department of Energy approvals) with a Zenodo DOI at: \href{https://doi.org}{https://doi.org}
 

### 2024-08-29 manuscript/supplement.tex
standalone supplement pdf

\documentclass{article}

\usepackage{multirow}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{inputenc}
\usepackage{geometry}
\usepackage{subfig}

\usepackage{ragged2e}  % for '\RaggedRight' macro
\usepackage{calc}      % for '\widthof' macro
\newcolumntype{P}[1]{>{\RaggedRight\hspace{0pt}}p{#1}}
\newcolumntype{L}{>{\RaggedRight}X}

\begin{document}

\noindent Table S1: All possible pairwise combinations of dependent and independent variables for the IRE case study (n=51). $R^2$ is the coefficient of determination of a linear model fit to the two covariates for the period of the abnormal release event (corresponding to 'during' in Fig. 2). Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter.

\input{table_s1}

\end{document}

### 2024-08-29 src/utils.py

@@ -92,6 +92,14 @@ def pdf_table(
         shell=True,
     )
 
+    if path_pdf == "tables/grid_all.pdf":
+        subprocess.call(
+            "pandoc --columns=10 mdtable.md -V fontsize=14pt -o "
+            + "manuscript/table_s1.tex",
+            # + " --pdf-engine-opt=-shell-escape",
+            shell=True,
+        )
+
     try:  # conda pdfcrop
         subprocess.check_call(
             "pdfcrop.pl " + path_pdf + " " + path_pdf,

### 2024-09-11 manuscript/manuscript.tex
espr submission edits

@@
 \renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.000/XXXXX}}
-\renewcommand{\undertitle}{This is a draft of a Manuscript in preparation for publication in Nature: Communications \\ \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
+\renewcommand{\undertitle}{This is a draft of a Manuscript intended for Environmental Science and Pollution Research \\ \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
 \renewcommand{\shorttitle}{Eddy covariance towers as sentinels}
@@
 
-\vspace{-4em}
+\vspace{-5em}
 
@@
 Because our approach is relatively simple and is resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration anomalies in flux tower data to ensure human health and industrial safety.
+
+\textbf{Keywords:} eddy covariance, data mining, industrial safety, vegetation response, radioactivity, photosynthesis
 
@@
 
-\section{Acknowledgements}
+\bibliographystyle{jsta}
+\bibliography{fluxnet}
+
+\section*{Statements and Declarations}
+
+\subsection*{Funding}
+
+We acknowledge funding for supporting the AmeriFlux data portal: U.S.
+Department of Energy Office of Science.
+The data used in this activity has been also funded by CarboEuropeIP (EU-FP6), IMECC (EU-FP6).
+This work was supported by Los Alamos National Laboratory (LDRD-20220062ER).
+
+\subsection*{Competing Interests}
+
+The authors have no relevant financial or non-financial interests to disclose.
+
+\subsection*{Author Contributions}
 
@@
 All authors provided interpretation of results.
-We acknowledge funding for supporting AmeriFlux data portal: U.S.
-Department of Energy Office of Science.
-The data used in this activity has been also funded by CarboEuropeIP (EU-FP6), IMECC (EU-FP6).
-This work was supported by Los Alamos National Laboratory (LDRD-20220062ER).
-
-\bibliographystyle{jsta}
-\bibliography{fluxnet}
 

### 2024-09-12 manuscript/manuscript.tex
espr submission re-edits

@@
 \author{
-	\href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Jemma Stachelek$^1$} \\
+	\href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Corresponding Author: Jemma Stachelek$^1$} \\
 	$^1$ High Performance Computing Division\\
@@
 
+\subsection*{Ethical Approval}
+
+Not applicable
+
+\subsection*{Consent to Participate}
+
+Not applicable
+
+\subsection*{Consent to Publish}
+
+Not applicable
+
+\subsection*{Availability of Data and Materials}
+
+No new data was produced in this study. All data used is available publicly from their respective sources and archived permanently with digital object identifiers.
+
 \end{document}

### 2024-09-12 manuscript/manuscript.tex
battle latex submission

@@
 	\And
-	E. Christi Thompson$^3$ \\
+	Elizabeth Christi Thompson$^3$ \\
 	$^3$ Computer, Computational, and Statistical Sciences Division\\
@@
 	\And
-	\textbf{L. Turin Dickman}$^2$ \\
+	\textbf{Lee Turin Dickman}$^2$ \\
 	$^2$ Earth and Environmental Sciences Division\\
@@
 	\centering
-	\includegraphics[width=16cm]{../figures/__map}
+	\includegraphics[width=16cm]{__map}
 	\caption{Map of flux tower locations for the IRE and Fukushima case studies (left and right respectively).
@@
 	\centering
-	\includegraphics[width=11cm]{../figures/__interaction_belon}
+	\includegraphics[width=11cm]{__interaction_belon}
 	\caption{Interaction plots at the BE-Lon site between air temperature ($C^{\circ}$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
@@
 	\centering
-	\includegraphics[width=16cm]{../figures/__rolling_fleurus}
+	\includegraphics[width=16cm]{__rolling_fleurus}
 	\caption{Time series plots (2004-2013) of rolling event detection analysis at the IRE site for the $CO_2$ and air temperature ($C^{\circ}$) variable pair.
@@
 	\centering
-	\includegraphics[width=16cm]{../figures/__hyperparameter_experiment}
+	\includegraphics[width=16cm]{__hyperparameter_experiment}
 	\caption{Sensitivity of event detection at the IRE site to different hyperparameter values (points) alongside potential relationship (solid line) and confidence interval (shaded region).}
@@
 	\centering
-	\includegraphics[width=16cm]{../figures/__rolling_fukushima}
+	\includegraphics[width=16cm]{__rolling_fukushima}
 	\caption{Time series plots (2010-2013) of rolling event detection analysis for the Fukushima disaster case study.
@@
 
-No new data was produced in this study. All data used is available publicly from their respective sources and archived permanently with digital object identifiers.
+No new data was produced in this study.
+All original data used is available publicly from their respective sources and archived permanently with restrictions on redistribution.
+These can be found at \url{http://www.europe-fluxdata.eu/}, \url{https://fluxnet.org/}, and \citep{cleverly2011alice}.
+
+Release of computer code and secondary data reuse that supports the results and analyses of the paper are pending a United State Department of Energy Review pursuant to \url{https://www.osti.gov/doecode/FAQs#how-do-i-announce-software-with-an-access-limitation-to-osti}.
 

### 2024-11-03 manuscript/manuscript.tex
super/sub script fixes

@@
 In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008.
-The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq 131I \citep{carle2010individual}.
+The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq $^{131}I$ \citep{carle2010individual}.
 For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg.
@@
 \subsection{Modeling Approach}
-We processed each data file by 1) excluding nighttime measurements based on photosynthetic photon flux density (PPFD > 100 µmol m-2 s-1) or net solar radiation (netrad > 0 W/m2) and 2) defining an “event period” following the publicly released date of the abnormal release.
+We processed each data file by 1) excluding nighttime measurements based on photosynthetic photon flux density (PPFD > 100 µmol m-2 s-1) or net solar radiation (netrad > 0 $W/m^2$) and 2) defining an “event period” following the publicly released date of the abnormal release.
 We focused on the following variables as potentially containing a signature of the release : carbon dioxide, carbon dioxide flux (fc), latent heat (le), and sensible heat flux (h).
@@
 This corresponds to a low effect size in Figure 2D.
-Note also how during and after the known abnormal release, the slopes of the relationships between CO2 and air temperature become less negative than before the event, with lower than expected CO2 at low air temperature (Figure 2A).
+Note also how during and after the known abnormal release, the slopes of the relationships between $CO_2$ and air temperature become less negative than before the event, with lower than expected $CO_2$ at low air temperature (Figure 2A).
 

### 2024-11-03 manuscript/manuscript.tex
subheader update

@@
 
-\subsection{Modeling Approach}
+\subsection{Statistical Approach}
 We processed each data file by 1) excluding nighttime measurements based on photosynthetic photon flux density (PPFD > 100 µmol m-2 s-1) or net solar radiation (netrad > 0 $W/m^2$) and 2) defining an “event period” following the publicly released date of the abnormal release.

### 2024-11-05 manuscript/manuscript.tex
stash edits

@@
 We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other eddy flux towers.
-While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, we expect that exposure may cause declines in photosynthetic $CO_2$ uptake.
-This may be expressed as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g.
+While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, prior studies support the idea that exposure may cause declines in photosynthetic $CO_2$ uptake.
+For example, \citet{gudkovEffectIonizingRadiation2019} detail how a wide range of exposure doses and durations can inhibit CO2 uptake as a result of enzymatic effects on the Calvin Cycle.
+For the present study, we lack leaf-level observations to directly establish inhibition and instead express it as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g.
 air temperature, humidity, precipitation, etc.) before, during, and after exposure.
@@
 
-For the IRE case study, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/).
+For both the IRE and Fukushima case studiesy, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) spanning multiple years.
+For the IRE case study, we targeted towers located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/).
 Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1A).
@@
 Additionally, we used wind direction data from the eddy flux towers to determine when or if any released material might plausibly reach a particular location.
-Each of these variables were recorded by the flux tower instrumentation at a 30 minute (min) interval barring any gaps due to missing data.
+All of the aforementioned data for each of these variables obtained from flux tower instrumentation recorded at a 30 minute (min) interval barring any gaps due to missing data.
 
@@
 In these cases, it may be possible to continuously mine flux tower records (as they become available) for events that exceed some threshold identified in a more comprehensive cross-site, cross-event benchmarking study.
+The delay between data collection and availability to the general community is highly variable among sites and investigators.
+In the best case scenario, for actively monitored sites with responsive investigators, this can be as little as 6 months.
+As a result, operationalization of our approach might target post-hoc attribution rather than real-time monitoring applications.
 
-More generally, we show that flux tower records have value for this type of data mining despite the apparent noisiness of the data \citep{fratiniEddyCovarianceFlux2018}.
+More generally, we show that flux tower records have value for this type of data mining despite the apparent noisiness of the data \citep[e.g., ][]{fratiniEddyCovarianceFlux2018}.
 We attribute the sensitivity of our approach to the fact that we fit successive interaction models to limited sections of the data record, and we leveraged bivariate relationships to constrain noisiness in individual data records.

### 2024-11-06 manuscript/manuscript.tex
stash edits

@@
 We specifically focused on short-term anomalies in the vegetation response data (on the order of days to weeks) due to either potential deposition of radioactive material or direct sensor effects rather than chronic long-term signals on the order of years to decades.
-In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008.
-The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq $^{131}I$ \citep{carle2010individual}.
-For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg.
-In the second case study, we further explored potential signals in flux tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}.
+In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium nominally occurring on August 23, 2008.
+This release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq $^{131}I$.
+The release went unreported (and possibly undetected) for several days.
+The first measurements in the vicinity of the facility were not taken until August 28th.
+These measurements found radioactivity levels of 5,000 Bq/kg in grass samples and led to a public health advisory for a 5km zone extending in a northeasterly direction from the facility \citep{carle2010individual}.
+Owing to the aforementioned measurement delays and the limited extent of the IRE release, we lack numeric estimates of the activity concentrations at tower locations.
+In the second case study, we further explored potential signals in flux tower data related to the Fukushima Nuclear Disaster occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere.
+An atmospheric dispersion study by \citet{meszarosPredictabilityDispersionFukushimaderived2016} estimates that our selected tower locations experienced deposition of particulate matter in the range of 0.44 - 31.08 $mBq/m^{3}$.
+
 We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other eddy flux towers.
@@
 In order to flag event detections, we defined several adjustable hyperparameters to account for uncertainty surrounding atmospheric transport and the timing of any possible responses to material deposition.
-The first hyperparameter we defined, wind\_tolerance, reflects the range around the bearing from the facility to the tower location which counts as towards the tower.
-In the base default case, we set wind\_tolerance to 10 degrees, but we explored setting it to smaller values down to 5 degrees and larger values up to 20 degrees.
-The second hyperparameter we defined, event\_quantile\_effect, handles the unknown travel time of any materials to be deposited via an abnormal release or what, if any, delays exist between deposition and vegetation response.
-Rather than simply selecting the maximum value of $\beta$3 during a given window or the value at the exact time of the event, the value of event\_quantile\_effect is set to correspond with a specific quantile of all $\beta$3 values during the event window.
+The first hyperparameter we defined, \emph{wind\_tolerance}, reflects the range around the bearing from the facility to the tower location which we count as towards the tower.
+In the base default case, we set \emph{wind\_tolerance} to 10 degrees, but we explored setting it to smaller values down to 5 degrees and larger values up to 20 degrees.
+The second hyperparameter we defined, \emph{event\_quantile\_effect}, handles the unknown travel time of any materials to be deposited via an abnormal release or what, if any, delays exist between deposition and vegetation response.
+Rather than simply selecting the maximum value of $\beta$3 during a given window or the value at the exact time of the event, the value of \emph{event\_quantile\_effect} is set to correspond with a specific quantile of all $\beta$3 values during the event window.
 In the base default case, this was set at 0.9, but we explored values as low as 0.5.
@@
 
-% \section{Data Availability Statement}
-
-% All code associated with this manuscript will be made available by the time of publication (or earlier pending the conclusion of US Department of Energy approvals) with a Zenodo DOI at: \href{https://doi.org}{https://doi.org}
-
 \bibliographystyle{jsta}
@@
 
-Release of computer code and secondary data reuse that supports the results and analyses of the paper are pending a United State Department of Energy Review pursuant to \url{https://www.osti.gov/doecode/FAQs#how-do-i-announce-software-with-an-access-limitation-to-osti}.
+Computer code and secondary data reuse that supports the results and analyses of the paper are available at: \href{https://doi.org/10.5281/zenodo.13845254}{https://doi.org/10.5281/zenodo.13845254}.
 

### 2024-11-07 manuscript/manuscript.tex
stash text edits

@@
 In the base default case, this was set at 0.9, but we explored values as low as 0.5.
-All statistical analyses were carried out using the statsmodels Python package (Seabold and Perktold, 2009).
-Our processing scripts are openly available at [doi link], and we refer readers to the original providers for data access.
+Taken together, here is the overall logic we used to flag an event detection: consider a scenario where the effect size value (corresponding to the \emph{event\_quantile\_effect} of the distribution of effect sizes during the \emph{n\_days} window length) during a known event is X at a tower which is Z bearing from the known release location.
+An event detection would be flagged during any period where the effect size is $\geq$ X and the wind direction is Z $\pm$ \emph{wind\_tolerance}.
+
+All statistical analyses in the present study were carried out using the statsmodels Python package (Seabold and Perktold, 2009).
+Our processing scripts are openly available at \href{https://doi.org/10.5281/zenodo.13845254}{https://doi.org/10.5281/zenodo.13845254}, and we refer readers to the original providers for data access.
 
@@
 Conversely, at the northerly BE-Bra site, the algorithm was not able to identify the known release event distinct from background noise.
-This can be seen in the rate of event detections ($\alpha$) at each site, whereby it was less than 1\% at the BE-Lon and BE-Vie sites, but was greater than 2\% at the BE-Bra site.
-It is notable that this ability to uniquely detect the known release along with a plausible number of additional event detections corresponds with the fraction of the abnormal release period when the wind was pointed from the release location towards the respective tower (using the default base case wind tolerance of 10 deg) which was much lower at the BE-Bra site (5\%) compared to the BE-Lon and BE-Vie sites (34\% Table 1).
+This can be seen in the rate of event detections ($\alpha$) at each site, calculated as the number of event detections divided by the total number of rolling windows in the period of record.
+This rate was less than 1\% at the BE-Lon and BE-Vie sites, but was greater than 2\% at the BE-Bra site.
+It is notable that this ability to uniquely detect the known release along with a plausible number of additional event detections corresponds with the fraction of the abnormal release period when the wind was pointed from the release location towards the respective tower.
+The fraction of the period of record where the wind was pointed from the facility toward the flux towers was much lower at the BE-Bra site (5\%) compared to the BE-Lon and BE-Vie sites (34\% Table 1).
 
@@
 We did, however, observe that the event detection was related to the window length and event quantile parameters as shown by the slope of the line in Figure 4.
-Although increasing values of the event quantile parameter would hypothetically increase event detection rates, we do not see value in setting this parameter below the median.
-Similarly, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.
+Although increasing values of the event quantile parameter above the limit of our tests ($>$ 0.9) would hypothetically increase event detection rates, we do not show this due to the noisiness of the extreme quantiles of effect size distributions.
+Additionally, we do not show tests setting this parameter below the median because it would lead to  weak sensitivity to abnormal events.
+In a similar fashion as the event quantile parameter, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not show tests setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts showing limited deposition at such long time scales \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.
 

### 2024-11-08 manuscript/manuscript.tex
stash text edits

@@
 Figure 2 provides a visual example of the effect size framing in our analysis.
-Note how the apparent interaction effect between time periods around the abnormal release event in Figure 2A (before, during, after) translates to a relatively high effect size (Figure 2C) whereas in Figure 2B at an arbitrary time point, there is no apparent interaction effect.
+Note how the apparent interaction effect between time periods around the abnormal release event in Figure 2A (i.e., different slopes for before, during, after) translates to a relatively high effect size (Figure 2C) whereas in Figure 2B at an arbitrary time point, there is no apparent interaction effect.
 This corresponds to a low effect size in Figure 2D.
@@
 In these cases, it may be possible to continuously mine flux tower records (as they become available) for events that exceed some threshold identified in a more comprehensive cross-site, cross-event benchmarking study.
-The delay between data collection and availability to the general community is highly variable among sites and investigators.
+The delay between data collection and availability to the general community via data repositories (e.g., Fluxnet) is highly variable among sites and investigators.
 In the best case scenario, for actively monitored sites with responsive investigators, this can be as little as 6 months.

### 2024-11-11 src/utils.py
add fukushima table to supplement

@@ -100,6 +100,14 @@ def pdf_table(
             shell=True,
         )
 
+    if path_pdf == "tables/grid_all_fukushima.pdf":
+        subprocess.call(
+            "pandoc --columns=10 mdtable.md -V fontsize=14pt -o "
+            + "manuscript/table_s1_fukushima.tex",
+            # + " --pdf-engine-opt=-shell-escape",
+            shell=True,
+        )
+
     try:  # conda pdfcrop
         subprocess.check_call(
             "pdfcrop.pl " + path_pdf + " " + path_pdf,

### 2024-11-11 tables/grid_all.py

@@ -22,6 +22,7 @@ def process_grid(i):
 # values here are different than the overview table because here they're relative to the POR
 # in the overview table they're limited to only during the event
 
+# # IRE case study
 flist = ["data/grid_be-lon_7.csv", "data/grid_be-vie_7.csv", "data/grid_be-bra_7.csv"]
 grid_list = [process_grid(i) for i in range(len(flist))]
 
@@ -29,6 +30,19 @@ utils.pdf_table(
     list(itertools.chain(*grid_list)),
     "",
     "tables/grid_all.pdf",
+    ["Dep", "Indep", "R2", "Effect", "Site"],
+    render_only=False,
+)
+
+# # Fukushima case study
+
+flist = ["data/grid_oz-mul_7.csv", "data/grid_us-gle_7.csv", "data/grid_us-wrc_7.csv"]
+grid_list = [process_grid(i) for i in range(len(flist))]
+
+utils.pdf_table(
+    list(itertools.chain(*grid_list)),
+    "",
+    "tables/grid_all_fukushima.pdf",
     ["Dep", "Indep", "R2", "Effect", "Site"],
     render_only=False,
 )

### 2024-11-11 tables/grid_all.py
handle extra column

@@ -13,6 +13,11 @@ def process_grid(i):
     dt["site"] = utils.lower_to_mixed_case(flist[i].split("_")[1])
     dt = dt[dt["r2"] > 0].sort_values("r2", ascending=False)
     dt["fquant"] = dt["fquant"] * 100
+
+    dt_names = [x for x in dt.columns]
+    if "n_days" in dt_names:
+        dt = dt[list(itertools.compress(dt_names, [x != "n_days" for x in dt_names]))]
+
     dt = dt.values.tolist()
     # dt = dt + [SEPARATING_LINE]
     return dt

### 2024-11-12 manuscript/supplement.tex
add fukushima table to supplement
add definitions to supplement
add a fukushima version of fig 2

@@
 \usepackage{subfig}
+\usepackage{float}
 
@@
 \begin{document}
+\title{Supplmentary material to: "Eddy covariance towers as sentinels of abnormal radioactive material releases"}
+\date{}
+\maketitle
+
+\makeatletter 
+\renewcommand{\thefigure}{S\@arabic\c@figure}
+\makeatother
+
+\begin{figure}
+	\centering
+	\includegraphics[width=11cm]{__interaction_uswrc}
+	\caption{Interaction plots at the US-Wrc site between latent heat and relative humidity for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
+	\label{fig:interaction_fukushima}
+\end{figure}
+
+\noindent Table S0: Variable definitions for Tables S1 and S2.
+\input{table_s0}
 
@@
 
+\noindent Table S2: All possible pairwise combinations of dependent and independent variables for the Fukushima case study (n=51).
+$R^2$ is the coefficient of determination of a linear model fit to the two covariates for the period of the abnormal release event (corresponding to 'during' in Fig.
+2).
+Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter.
+
+\input{table_s1_fukushima}
+
 \end{document}

### 2024-11-12 figures/interaction.py

@@ -23,7 +23,7 @@ def hue_regplot(data, x, y, hue, palette=None, **kwargs):
     levels = data[hue].unique()
 
     if palette is None:
-        default_colors = get_cmap("tab10")
+        default_colors = get_cmap("Dark2")
         palette = {k: default_colors(i) for i, k in enumerate(levels)}
 
     for key in levels:
@@ -47,6 +47,8 @@ n_days = 7
 event_quantile = 0.9
 var_dep = "co2"
 var_idep = "ta"
+ylab = "$CO_2$ (ppm)"
+xlab = r"Temperature ($^{\circ}$C)"
 # var_dep = "fc"
 # var_idep = "netrad"
 
@@ -82,33 +84,6 @@ dt, dt_select = rolling.preprocess_dt(path_in, dep_cols, indep_cols)
 dt = janitor.remove_empty(dt)
 dt_event = rolling.define_period(dt_select, n_days=n_days, date_event=date_event)
 
-plt.close()
-fig, axs = plt.subplots(2, 2)
-ax1, ax2, ax3, ax4 = axs[0, 0], axs[0, 1], axs[1, 0], axs[1, 1]
-gg1 = hue_regplot(data=dt_event, x="ta", y="co2", hue="period", ax=ax1, ci=None)
-colors = sns.color_palette(n_colors=3).as_hex()
-legend_elements = [
-    Line2D([0], [0], color=colors[0], lw=4, label="Before"),
-    Line2D([0], [0], color=colors[1], lw=4, label="During"),
-    Line2D([0], [0], color=colors[2], lw=4, label="After"),
-]
-ax1.legend(handles=legend_elements, loc="upper right")
-ax1.set_ylabel("$CO_2$ (ppm)")
-ax1.set_xlabel(r"Temperature ($^{\circ}$C)")
-ax1.set_title("A.", pad=0, x=0.07, y=0.9)
-#
-gg2 = sns.histplot(
-    x=pfdist["fdist"],
-    element="step",
-    edgecolor="gray",
-    color="gray",
-    ax=ax2,
-)
-gg2.axvline(fevent["fevent"].values[0], color="orange")
-ax2.set_xlabel("Effect size")
-ax2.set_title("C.", pad=0, x=0.07, y=0.9)
-#
-
 # limit ts_start to the same month as date_event
 pfdist["month"] = [x.month for x in pd.to_datetime(pfdist["timestamp"])]
 # avoid 2009 because of missing data
[…]

### 2024-11-12 figures/interaction.py
use open circles per reviewer

@@ -29,7 +29,12 @@ def hue_regplot(data, x, y, hue, palette=None, **kwargs):
     for key in levels:
         regplots.append(
             sns.regplot(
-                x=x, y=y, data=data[data[hue] == key], color=palette[key], **kwargs
+                x=x,
+                y=y,
+                data=data[data[hue] == key],
+                color=palette[key],
+                marker="$\circ$",
+                **kwargs
             )
         )
 

### 2024-11-12 manuscript/manuscript.tex
organizing and discussing supplement figures

@@
 
-To begin our analysis, we screened all pairwise, linear relationships between the aforementioned independent and dependent variables using a linear regression fit between each pair for the entire period of record (n=51).
+To begin our analysis, we screened all pairwise, linear relationships between the aforementioned independent and dependent variables using a linear regression fit between each pair for the entire period of record (Table S1, n=51).
 Variable pairs were excluded from further analysis if the overall coefficient of determination ($R^2$) was less than 0.1.
@@
 
-For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51, Table S1), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model ($R^2$ > 0.05) and a strong event effect size (Table 1).
+For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51, Table S1), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model and a strong event effect size (Table 1).
+Note that values in Table S1 are computed relative to the entirety of the data record for variable pair screening purposes, while the values in Table 1 are computed over the known abnormal event period for evaluating event identifiability.
 Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites.
@@
 Note also how during and after the known abnormal release, the slopes of the relationships between $CO_2$ and air temperature become less negative than before the event, with lower than expected $CO_2$ at low air temperature (Figure 2A).
+A similar pattern is apparent for the BE-Bra and BE-Vie sites (Figure S1, S2).
 

### 2024-11-12 manuscript/supplement.tex
organizing and discussing supplement figures

@@
 
+\pagenumbering{gobble}
+
 \begin{document}
@@
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
 	\includegraphics[width=11cm]{__interaction_uswrc}
-	\caption{Interaction plots at the US-Wrc site between latent heat and relative humidity for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+	\caption{Interaction plots at the BE-Bra site between air temperature ($C^{\circ}$) and $CO_2$ (ppm)for the period of the known abnormal release (A) and an arbitrary non-release period (B).
 Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
-	\label{fig:interaction_fukushima}
+	\label{fig:interaction_bebra}
 \end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=11cm]{__interaction_bevie}
+	\caption{Interaction plots at the BE-Vie site between air temperature ($C^{\circ}$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
+	\label{fig:interaction_bevie}
+\end{figure}
+
+% \begin{figure}[H]
+% 	\centering
+% 	\includegraphics[width=11cm]{__interaction_uswrc}
+% 	\caption{Interaction plots at the US-Wrc site between latent heat and relative humidity for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
+% 	\label{fig:interaction_fukushima}
+% \end{figure}
+
+\newpage
 
@@
 \noindent Table S1: All possible pairwise combinations of dependent and independent variables for the IRE case study (n=51).
-$R^2$ is the coefficient of determination of a linear model fit to the two covariates for the period of the abnormal release event (corresponding to 'during' in Fig.
+$R^2$ is the coefficient of determination of a linear model fit to the two covariates for the entirety of the period of record (not specific to any window in Fig.
 2).
-Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter.
+Effect is the event effect size of the interaction term 'during' the known release period matching the 100\% quantile of the distribution.
 
@@
 
-\noindent Table S2: All possible pairwise combinations of dependent and independent variables for the Fukushima case study (n=51).
-$R^2$ is the coefficient of determination of a linear model fit to the two covariates for the period of the abnormal release event (corresponding to 'during' in Fig.
+\noindent Table S2: All possible pairwise combinations of dependent and independent variables for the Fukushima case study (n=51) for screening of candidate variable pairs.
+$R^2$ is the coefficient of determination of a linear model fit to the two covariates for the entirety of the period of record (not specific to any window in Fig.
 2).
-Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter.
+Effect is the event effect size of the interaction term 'during' the known release period matching the 100\% quantile of the distribution.
 

### 2024-11-12 figures/interaction.py

@@ -41,6 +41,76 @@ def hue_regplot(data, x, y, hue, palette=None, **kwargs):
     return regplots
 
 
+def interaction_plot(
+    dt_event,
+    var_idep,
+    var_dep,
+    ylab,
+    xlab,
+    pfdist,
+    fevent,
+    dt_select,
+    n_days,
+    ts_start,
+    out_path,
+):
+    plt.close()
+    _, axs = plt.subplots(2, 2)
+    ax1, ax2, ax3, ax4 = axs[0, 0], axs[0, 1], axs[1, 0], axs[1, 1]
+    #
+    gg1 = hue_regplot(
+        data=dt_event, x=var_idep, y=var_dep, hue="period", ax=ax1, ci=None
+    )
+    colors = sns.color_palette(palette="Dark2", n_colors=3).as_hex()
+    legend_elements = [
+        Line2D([0], [0], color=colors[0], lw=4, label="Before"),
+        Line2D([0], [0], color=colors[1], lw=4, label="During"),
+        Line2D([0], [0], color=colors[2], lw=4, label="After"),
+    ]
+    ax1.legend(handles=legend_elements, loc="upper right")
+    ax1.set_ylabel(ylab)
+    ax1.set_xlabel(xlab)
+    ax1.set_title("A.", pad=0, x=0.07, y=0.9)
+    #
+    gg2 = sns.histplot(
+        x=pfdist["fdist"],
+        element="step",
+        edgecolor="gray",
+        color="gray",
+        ax=ax2,
+    )
+    gg2.axvline(fevent["fevent"].values[0], color="orange")
+    gg2.set_xlim(0, 150)
+    ax2.set_xlabel("Effect size")
+    ax2.set_title("C.", pad=0, x=0.07, y=0.9)
+    #
+    dt_event_weak = rolling.define_period(dt_select, n_days=n_days, date_event=ts_start)
+    gg3 = hue_regplot(
+        data=dt_event_weak, x="ta", y="co2", hue="period", ax=ax3, ci=None
+    )
+    ax3.set_ylabel(ylab)
+    ax3.set_xlabel(xlab)
+    ax3.set_title("B.", pad=0, x=0.07, y=0.9)
+    #
+    # stats.percentileofscore(pfdist["fdist"], fevent["fevent"], nan_policy="omit") / 100
+    gg4 = sns.histplot(
+        x=pfdist["fdist"],
+        element="step",
[…]

### 2024-11-12 manuscript/manuscript.tex
stash edits

@@
 
-For both the IRE and Fukushima case studiesy, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) spanning multiple years.
+For both the IRE and Fukushima case studies, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) spanning multiple years.
 For the IRE case study, we targeted towers located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/).
@@
 For the IRE case study, among all possible pairwise combinations of dependent and independent variables (n=51, Table S1), we found that only the $CO_2$ versus air temperature comparison had any measure of predictability from the model and a strong event effect size (Table 1).
-Note that values in Table S1 are computed relative to the entirety of the data record for variable pair screening purposes, while the values in Table 1 are computed over the known abnormal event period for evaluating event identifiability.
+Note that values in Table S1 are computed relative to the entirety of the data record for variable pair screening purposes, while the values in Table 1 are computed over the known abnormal event period for event identifiability purposes.
 Although other variable pairs such as sensible heat flux and net radiation had a stronger effect size for the BE-Lon site, they had a weak linear relationship across the other sites.
@@
 Although increasing values of the event quantile parameter above the limit of our tests ($>$ 0.9) would hypothetically increase event detection rates, we do not show this due to the noisiness of the extreme quantiles of effect size distributions.
-Additionally, we do not show tests setting this parameter below the median because it would lead to  weak sensitivity to abnormal events.
+Additionally, we do not show tests setting this parameter below the median because it would lead to weak sensitivity to abnormal event detection.
 In a similar fashion as the event quantile parameter, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not show tests setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts showing limited deposition at such long time scales \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.
@@
 We attribute the sensitivity of our approach to the fact that we fit successive interaction models to limited sections of the data record, and we leveraged bivariate relationships to constrain noisiness in individual data records.
-Our approach was aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers.
+This strategy of using only limited sections of the data is particularly important because it can avoid the situation where a well-defined linear relationship becomes non-linear and obscured at longer timescales.
+Despite the apparent success of our approach, it is not a perfect solution because some non-linearity remains.
+Beyond the noise inherent to most environmental data, this is likely one of the principal causes of the relatively modest $R^2$ values we observed in Table 1.
+These are not unusally low however as $R^2$ values between 0.35 and 0.7 are typical for environmental data \citep{murrayMethodsQuantifyVariable2009}.
+A fruitful area of future research would be exploring physical justifications for linearizing these relationships, which would likely lead to more precise signal recovery and event detection.
+Our approach was further aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers.
 This is because although such systematic errors may bias the magnitude of overall flux values, which would be a problem for typical uses of flux towers that deal directly in the units of the measured data, they do not reduce the overall confidence in individual values \citep{langfordEddycovarianceDataLow2015}, which for our purposes might represent an abnormal release signal.
-The ability of our approach to identify a suitable release signature does likely depend on the quality and completeness of the underlying data.
+As a result, the ability of our approach to identify a suitable release signature does depend on the quality and completeness of the underlying data.
 

### 2024-11-12 figures/interaction.py

@@ -80,7 +80,7 @@ def interaction_plot(
         ax=ax2,
     )
     gg2.axvline(fevent["fevent"].values[0], color="orange")
-    gg2.set_xlim(0, 150)
+    gg2.set_xlim(-2, 150)
     ax2.set_xlabel("Effect size")
     ax2.set_title("C.", pad=0, x=0.07, y=0.9)
     #
@@ -101,7 +101,7 @@ def interaction_plot(
         ax=ax4,
     )
     gg4.axvline(np.nanmin(pfdist["fdist"]), color="orange")
-    gg4.set_xlim(0, 150)
+    gg4.set_xlim(-2, 150)
     ax4.set_xlabel("Effect size")
     ax4.set_title("D.", pad=0, x=0.07, y=0.9)
     #

### 2024-11-19 manuscript/manuscript.tex
text edits

@@
 Our utilization of flux tower records for event discrimination stands in contrast to the typical use of data from the towers for direct carbon balance accounting.
-Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange \citep{baldocchiInterannualVariabilityNet2018}, our use treats each variable as potentially including an indirect signal (i.e.
-an anomaly from normal vegetation behavior) of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc).
+Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange \citep{baldocchiInterannualVariabilityNet2018}, our use treats each variable as potentially including an indirect signal (i.e., an anomaly from normal vegetation behavior or expected carbon or water vapor fluxes) of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc).
 In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation in a known way \citep{nobelPhysicochemicalEnvironmentalPlant2020}.
@@
 For example, \citet{gudkovEffectIonizingRadiation2019} detail how a wide range of exposure doses and durations can inhibit CO2 uptake as a result of enzymatic effects on the Calvin Cycle.
-For the present study, we lack leaf-level observations to directly establish inhibition and instead express it as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g.
+For the present study, we lack leaf-level observations to directly establish inhibition and instead express it indirectly as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g.
 air temperature, humidity, precipitation, etc.) before, during, and after exposure.
+A consequence of this indirect approach is that we cannot directly separate material deposition responses coming from vegetation apart from possible deposition interference with flux tower sensors.
 
@@
 In the best case scenario, for actively monitored sites with responsive investigators, this can be as little as 6 months.
-As a result, operationalization of our approach might target post-hoc attribution rather than real-time monitoring applications.
+As a result, operationalization of our approach might target post-hoc attribution rather than real-time monitoring applications unless release protocols of flux data are increased in frequency.
 
@@
 These are not unusally low however as $R^2$ values between 0.35 and 0.7 are typical for environmental data \citep{murrayMethodsQuantifyVariable2009}.
-A fruitful area of future research would be exploring physical justifications for linearizing these relationships, which would likely lead to more precise signal recovery and event detection.
+A fruitful area of future research would be exploring physical justifications for including non-linear relationships, which would likely lead to more precise signal recovery and event detection.
 Our approach was further aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers.

### 2025-02-19 manuscript/manuscript.tex
acceptance formatting,typo fixes

@@
 \author{
+	% 1
 	\href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Corresponding Author: Jemma Stachelek$^1$} \\
@@
 	\And
+	% 4
+	\textbf{Lee Turin Dickman}$^2$ \\
+	$^2$ Earth and Environmental Sciences Division\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\	
+	\And
+	% 2
 	Vachel A. Kraklow$^2$ \\
@@
 	\And
+	% 5
+	\textbf{Emily Casleton}$^3$ \\
+	$^3$ Computer, Computational, and Statistical Sciences Division\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\	
+	\And
+	% 3
 	Elizabeth Christi Thompson$^3$ \\
@@
 	\And
-	\textbf{Lee Turin Dickman}$^2$ \\
-	$^2$ Earth and Environmental Sciences Division\\
-	Los Alamos National Laboratory\\
-	Los Alamos, NM, USA, 87544 \\
-	\And
-	\textbf{Emily Casleton}$^3$ \\
-	$^3$ Computer, Computational, and Statistical Sciences Division\\
-	Los Alamos National Laboratory\\
-	Los Alamos, NM, USA, 87544 \\
-	\And
+	% 6
 	\textbf{Sanna Sevanto}$^2$ \\
@@
 	\includegraphics[width=11cm]{__interaction_belon}
-	\caption{Interaction plots at the BE-Lon site between air temperature ($C^{\circ}$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+	\caption{Interaction plots at the BE-Lon site between air temperature ($^{\circ}C$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
 Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
@@
 	\includegraphics[width=16cm]{__rolling_fleurus}
-	\caption{Time series plots (2004-2013) of rolling event detection analysis at the IRE site for the $CO_2$ and air temperature ($C^{\circ}$) variable pair.
+	\caption{Time series plots (2004-2013) of rolling event detection analysis at the IRE site for the $CO_2$ and air temperature ($^{\circ}C$) variable pair.
 Event detection lines (orange), interaction effect size (blue), and event effect size (dashed black line) are shown in the top panels.
@@
 Such information may help disentangle potential interference of atmospheric contaminants with infrared measurement of $CO_2$ from declines in photosynthetic $CO_2$ uptake of contaminated vegetation given our observation that during and after the known abnormal release, the slopes of the relationships between $CO_2$ and air temperature become less negative than before the event, with lower than expected $CO_2$ at low air temperature.
-Although the chemical makeup of possible atmospheric contaminants is unknown, interference by iodine itself is unlikely to affect measurements of $CO_2$ given that the absorption spectra of iodine peaks at much shorter wavelengths around 10-7 m \citep{haynes2016crc} compared to that of $CO_2$ which peaks around 1700-2100 10-9 m (LICOR Biosciences Inc., Lincoln, NE).
+Although the chemical makeup of possible atmospheric contaminants is unknown, interference by iodine itself is unlikely to affect measurements of $CO_2$ given that the absorption spectra of iodine peaks at much shorter wavelengths around $10^{-7}$ m \citep{haynes2016crc} compared to that of $CO_2$ which peaks around 1700-2100 $10^{-9}$ m (LICOR Biosciences Inc., Lincoln, NE).
 

### 2025-02-19 manuscript/supplement.tex
acceptance formatting,typo fixes

@@
 \begin{document}
-\title{Supplmentary material to: "Eddy covariance towers as sentinels of abnormal radioactive material releases"}
+\title{Supplementary material to: "Eddy covariance towers as sentinels of abnormal radioactive material releases"}
 \date{}
@@
 	\includegraphics[width=11cm]{__interaction_uswrc}
-	\caption{Interaction plots at the BE-Bra site between air temperature ($C^{\circ}$) and $CO_2$ (ppm)for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+	\caption{Interaction plots at the BE-Bra site between air temperature ($^{\circ}C$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
 Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}
@@
 	\includegraphics[width=11cm]{__interaction_bevie}
-	\caption{Interaction plots at the BE-Vie site between air temperature ($C^{\circ}$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
+	\caption{Interaction plots at the BE-Vie site between air temperature ($^{\circ}C$) and $CO_2$ (ppm) for the period of the known abnormal release (A) and an arbitrary non-release period (B).
 Also shown is the distribution of all interaction effect sizes (gray) compared with the effect size of the time period shown in the corresponding left panels (orange, C, D).}

### 2025-02-19 manuscript/manuscript.tex
note acceptance

@@
 \renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.000/XXXXX}}
-\renewcommand{\undertitle}{This is a draft of a Manuscript intended for Environmental Science and Pollution Research \\ \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
+\renewcommand{\undertitle}{This is an Accepted Manuscript of an article in Environmental Science and Pollution Research \\ \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
 \renewcommand{\shorttitle}{Eddy covariance towers as sentinels}

### 2025-02-24 manuscript/manuscript.tex
add paper doi

--- 2025-02-19 manuscript/manuscript.tex
@@
 % Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.000/XXXXX}}
-\renewcommand{\undertitle}{This is a draft of a Manuscript intended for Environmental Science and Pollution Research \\ \href{https://doi.org/10.1000/XXXXX}{DOI: 10.1000/XXXXX}}
+\renewcommand{\headeright}{\href{https://doi.org/10.1007/s11356-025-36171-3}{DOI: 10.1007/s11356-025-36171-3}}
+\renewcommand{\undertitle}{This is an Accepted Manuscript of an article in Environmental Science and Pollution Research \\ \href{https://doi.org/10.1000/10.1007/s11356-025-36171-3}{DOI: 10.1000/10.1007/s11356-025-36171-3}}
 \renewcommand{\shorttitle}{Eddy covariance towers as sentinels}

### 2025-02-24 scripts/00_nk_test_dates.py
cleanup

@@ -1,27 +1,0 @@
-import numpy as np
-import pandas as pd
-
-# Oct 2006, First underground test that released gases
-# May 2009, Second underground test
-# Feb 2013, Third underground test, gases detected in Japan
-# Jan 6th, 2016 H2 test about 6-10kt underground
-# Sept, 2016 H2 underground test
-# Sept 2017 H2 underground test, largest of all
-
-nktests = pd.DataFrame(
-    np.array(
-        [
-            ["2006-10-15"],
-            ["2009-05-15"],
-            ["2013-02-15"],
-            ["2016-01-06"],
-            ["2016-09-15"],
-            ["2017-09-15"],
-        ]
-    ),
-    columns=[
-        "date",
-    ],
-)
-nktests["date"] = pd.to_datetime(nktests["date"])
-nktests.to_csv("data/nktests.csv", index=False)
